### 2017-04-29 assignments/1/paper.tex
Initial commit.

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Assignment 2}

\author{Viktor Hansen, Asger Andersen \\ 
Department of Computer Science \\
University of Copenhagen}

\maketitle

\begin{abstract}
  This is the second weekly assignment for the Computability and Complexity course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{2.6b}
The complement of the language is $\left\{ w \; | \; w \in \left\{ a, b \right\}^* \right\} \setminus \left\{ a^nb^n \; | \; n \geq 0 \right\}$. Hence strings in the language $\left\{ \; a^nb^m \; | \; n,m \geq 0, n \neq m \right\} \subset \left\{ a^nb^m \; | \; n,m \geq 0 \right\}$ are in the complement while strings in $\left\{ \; a^nb^n \; | \; n \geq 0 \right\} \subset \left\{ a^nb^m \; | \; n,m \geq 0 \right\}$ are not, which needs to be reflected in the grammar:
\begin{equation*}
	\begin{split}
S &\longrightarrow aSb \; | \; aA \; | \; bB\\
A &\longrightarrow aA \; | \; \varepsilon \; | \; baD
	\end{split}
\quad\quad
	\begin{split}
B &\longrightarrow Bb \; | \; \varepsilon \; | \; aD\\
D &\longrightarrow aD \; | \; bD \; | \; \varepsilon
	\end{split}
\end{equation*}
We observe that at any given point a $b$ is followed by an $a$ in a derivation, the string remains in the complement, but not in $\left\{ a^nb^m \; | \; n,m \geq 0 \right\}$, regardless of any further derivation. This is captured in productions $A \rightarrow baD$ and $B \rightarrow aD$. The productions for $S$ enforce the requirement that no string in $\left\{ \; a^nb^n \; | \; n \geq 0 \right\}$ are in the complement. This is done by 'pumping' a string of the form $a \hdots a S b \hdots b$ and only allowing $S$ to become either $a$s, $b$s or a mix of either in which it is guaranteed that a $b$ is followed by an $a$ (productions $A \rightarrow baD$ and $B \rightarrow aD$). The production for $D$ derives strings in $\left\{a,b\right\}^{*}$, which is any string consisting solely of $a$s and $b$s.

\section*{2.6d}
The following grammar recognizes the language:
\begin{equation*}
	\begin{split}
F &\longrightarrow LPR
	\end{split}
\quad\quad
	\begin{split}
L &\longrightarrow LQ\# \; | \; \varepsilon \\
R &\longrightarrow \#QR \; | \; \varepsilon\\
Q &\longrightarrow aQ \; | \; bQ \; | \; \varepsilon
	\end{split}
\quad\quad
	\begin{split}
P &\longrightarrow aPa \; | \; bPb \; | \; P'\\
P' &\longrightarrow \#S\# \; | \; \# \; | \; a \; | \; b \; | \; \varepsilon\\
S &\longrightarrow Q \; | \; Q \# S \\
	\end{split}
\end{equation*}

The production for $Q$ recognizes the language $\left\{ a, b \right\}^*$, i.e. each substring $x_i$. The language requires that for some $i,j$, $x_i=x_j^{\mathcal{R}}$. In case that $i=j$, $x_i=x_j=x_i^{\mathcal{R}}$ is a palindrome, which is captured by the productions for $P$ and $P' \rightarrow a \; | \; b \; | \; \varepsilon$. When $x_{j}=x_{i+l}^{\mathcal{R}}$ for some integer $l \neq 0$, a string in the language $s = \left\{ \; \# x_1 \# x_2 \# \hdots \# x_k \# \; | \; k \geq 1,\text{ each  } x_i \in \left\{ a, b \right\}^* \; \right\}$ exists between $x_i$ and $x_j$, i.e. $x_{i} s x_{j}$, which is captured by the production $P' \rightarrow \#S\#$ and those for $S$. The productions for $L$ (and accordingly for $R$) optionally 'pad' the language with strings in the language $\left\{ \; x_1 \# x_2 \# \hdots \# x_k \# \; | \; k \geq 0,\text{ each  } x_i \in \left\{ a, b \right\}^* \; \right\}$ (and accordingly for $R$), allowing for $i$ and $j$ to be different from $1$ and $k$. The start production $F \rightarrow LPR$ yields the optional left 'padding', $L$, followed by the palindrome/reversed string part, followed by optional right 'padding'.

\section*{2.14}
First, the start symbol $S$ and its production is added to the grammar:
\begin{equation*}
	\begin{split}
		A &\rightarrow BAB \; | \; B \; | \; \varepsilon\\
		B &\rightarrow 00 \; | \; \varepsilon
	\end{split}
\quad\longrightarrow\quad
	\begin{split}
		\mathbf{S} &\rightarrow \mathbf{A}\\
    	A &\rightarrow BAB \; | \; B \; | \; \varepsilon\\
		B &\rightarrow 00 \; | \; \varepsilon
    \end{split}
\end{equation*}
Next, the $B \rightarrow \varepsilon$ and $A \rightarrow \varepsilon$ productions are eliminated, yielding:
\begin{equation*}
	\begin{split}
		S &\rightarrow A\\
    	A &\rightarrow BAB \; | \; B \; | \; \bm{\varepsilon}\\
		B &\rightarrow 00 \; | \; \bm{\varepsilon}
    \end{split}
\quad\longrightarrow\quad
	\begin{split}
		S &\rightarrow A \; | \; \bm{\varepsilon}\\
    	A &\rightarrow BAB \; | \; B \; | \; \mathbf{AB} \; | \; \mathbf{BA} \; | \; \mathbf{A} \; | \; \mathbf{BB}\\
		B &\rightarrow 00
    \end{split}
\end{equation*}
Now, unit rules $S \rightarrow A$, $A \rightarrow B$ and $A \rightarrow A$ are removed:
\begin{equation*}
	\begin{split}
		S &\rightarrow \mathbf{A} \; | \; \varepsilon\\
    	A &\rightarrow BAB \; | \; \mathbf{B} \; | \; AB \; | \; BA \; | \; \mathbf{A} \; | \; BB\\
		B &\rightarrow 00
    \end{split}
\quad\longrightarrow\quad
	\begin{split}
		S &\rightarrow \mathbf{BAB \; | \; 00 \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon}\\
    	A &\rightarrow BAB \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB\\
		B &\rightarrow 00
    \end{split}
\end{equation*}
Finally, the remaining rules are converted into Chomsky normal form by adding additional variables and rules to the grammar, yielding:
\begin{equation*}
	\begin{split}
		S &\rightarrow \mathbf{BA}B \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon\\
    	A &\rightarrow \mathbf{BA}B \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB\\
		B &\rightarrow 00
    \end{split}
\quad\longrightarrow\quad
	\begin{split}
		S &\rightarrow \mathbf{C}B \; | \; \mathbf{DD} \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon\\
    	A &\rightarrow \mathbf{C}B \; | \; \mathbf{DD} \; | \; AB \; | \; BA \; | \; BB\\
		\mathbf{C} &\rightarrow \mathbf{BA}\\
		\mathbf{B} &\rightarrow \mathbf{DD}\\
		\mathbf{D} &\rightarrow \mathbf{0}
    \end{split}
\end{equation*}

\section*{2.42}
\textbf{a)} Assume that the language $L_1 = \{0^n1^n0^n1^n| n\geq 0 \}$ is context-free. With this assumption, we get from the pumping lemma for context-free languages to that there exists a natural number p, such that all strings in $L_1$, which are at least p long, satisfy the three conditions in the lemma. But consider the string $s=0^p1^p0^p1^p$. Since $s$ is in $L_1$ and is longer than p, we know that $s$ satisfies the three conditions of the lemma. That is, we can write $s=uvxyz$ such that $|vy|>0$ and $|vxy|\leq p$ and $uv^ixy^iz$ is in $L_1$ for all $i\geq 0$. But since $|vxy|\leq p$, we have from the definition of $s$ that either $v$ and $y$ do not contain any 0's or they do not contain any 1's. Since $|vy|>0$, we have that at least one of $v$ or $y$ contains at least one 0 or 1. For that reason, when we pump $s$ to $uv^2xy^2z$ either the number of 0's will grow, while the number of 1's will not change, or the number of 1's will grow, while the number of 0's will not change. In both cases, the pumped string will have a different number of 0's and 1's. But all the strings in $L_1$ have the same number of 0's and 1's. Therefore, the pumped string is not in $L_1$. But from the pumping lemma, we know that $uv^2xy^2z$ is in $L_1$. This is a contradiction, which we have derived from our assumption that $L_1$ is context-free. This assumption is therefore false. $L_1$ is therefore not context-free.

\textbf{b)} Assume that the language $L_2 = \{t_1\#...\#t_k| k\geq 2,\  t_i\in \{ a,b \}^*,\ t_i = t_j \text{ for some } i\neq j \}$ is context-free. With this assumptions, we now get from the pumping lemma for context-free languages to that there exists a natural number p, such that all strings in $L_2$, which are at least p long, satisfy the three conditions in the lemma. But consider the string $s=a^pb^p \# a^pb^p$. Since $s$ is in $L_2$ and is longer than p, we know that $s$ satisfies the three conditions of the lemma. That is, we can write $s=uvxyz$ such that $|vy|>0$ and $|vxy|\leq p$ and $uv^ixy^iz$ is in $L_2$ for all $i\geq 0$.

All the ways of splitting up $s$ as $uvxyz$ can be categorized with these cases: 
\begin{enumerate}
\item None of $v$ or $y$ contains the $\#$ symbol:
\begin{enumerate}
\item $v$ and $y$ both contain symbols only from the same side of $\#$.
\item $v$ contains symbols only from the left side of $\#$, and $y$ contains symbols only from the right side of $\#$.
\end{enumerate} 
\item One of $v$ or $y$ contains the $\#$ symbol.
\end{enumerate}

Let us now argue why for each case we cannot find $v$ and $y$ such that $s=uvxyz$ can be pumped. \begin{enumerate}
\item
\begin{enumerate}
\item Since none of $v$ or $y$ contain the $\#$, then $uv^2xy^2z$ still only contains one $\#$. Since $v$ and $y$ cannot both be empty strings and they both only contains symbols from one side of the $\#$, then the pumping only effects one side of the $\#$. Therefore $uv^2xy^2z$ will have the form $t_1\#t_2$, where $t_1\neq t_2$. This string is not in $L_2$.
\item Since none of $v$ or $y$ contain the $\#$, then $uv^2xy^2z$ still only contains one $\#$. Since $v$ and $y$ contain symbols from different sides of $\#$ and none of the contains $\#$, then $x$ must contain $\#$. Since $|vxy|\leq p$, then we can conclude that $v$ cannot contain other symbols than $b's$, and $y$ cannot contain other symbols than $a's$. Since $v$ and $y$ cannot both be empty strings, then we can now conclude that $uv^2xy^2z$ will have the form $t_1\#t_2$, where $t_1\neq t_2$. This string is not in $L_2$.
\end{enumerate}
\item Since $v$ or $y$ contains the $\#$, then $uv^0xy^0z$ will not have any $\#$. In other words, it will just have the form $t_1$, where $t_1$ is some string in $\{a,b\}^*$. This string is not in $L_2$.
\end{enumerate}

\section*{2.54}
We consider the possible ways to divide the string $s=1^0\#1^1\#1^2\# \hdots \#1^p$ according to the pumping lemma for CFGs. Obviously $s$ is in the language and the pumping lemma states that since $\left| s \right| \geq p$, it can be divided it into parts $uvxyz$, st. $\left| vxy \right| \leq p$, $\left| vy \right| \geq 0$ and $uv^ixy^iz$ is in the language for $i \geq 0$.

Choosing either $v$ or $y$ as $\#$, pumping it up would yield more than one empty sequence between the sharps in $s$, and the resulting string would not be in the language. Choosing either $v$ or $u$ as any substring containing at least one sharp would also lead to a contradiction when pumping up (or down), by similar reasoning. Choosing either $v$ or $u$ as all ones will lead to a contradiction when pumping down, since this will result in a sequence of ones that already exists elsewhere. Specifically, choosing a length $k \leq p$ all-ones substring of any all-ones substring and pumping down will yield another all-ones substring of size at most $p-k$. But all such strings $1^{p-k}$ for $k \leq p$ are already a substring between sharps in $s$, leading to the contradiction and hence the grammar is not context free.

%\begin{figure}[!hbt]
%\center
%\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,semithick]
%  \tikzstyle{every state}=[fill=red,draw=none,text=white]

%  \node[initial,state] (A)                    {$q_a$};
%  \node[state]         (B) [above right of=A] {$q_b$};
%  \node[state]         (D) [below right of=A] {$q_d$};
%  \node[state]         (C) [below right of=B] {$q_c$};
%  \node[state]         (E) [below of=D]       {$q_e$};

%  \path (A) edge              node {0,1,L} (B)
%            edge              node {1,1,R} (C)
%        (B) edge [loop above] node {1,1,L} (B)
%            edge              node {0,1,L} (C)
%        (C) edge              node {0,1,L} (D)
%            edge [bend left]  node {1,0,R} (E)
%        (D) edge [loop below] node {1,1,R} (D)
%            edge              node {0,1,R} (A)
%        (E) edge [bend left]  node {1,0,R} (A);
%\end{tikzpicture}
%  \caption{Output of running the program, showing the values produced by each %step.\label{lst:output}}
%\end{figure}

\section*{3.16}
The book's solution to 3.16a) is a quite high-level description of how to combine a TM $M_1$ deciding a language $L_1$ and a TM $M_2$ deciding the language $L_2$ into a new TM $M$ that decides the union of the two languages. Our solution will follow this high-level style (that is, not specifying the details of the 7-tuples that describes the different turing-machines). We will be more careful than the book in the sense that we use non-determinism to make it a little more explicit how the new machine M makes sure that one of the machines $M_i$ does not end up working on something that it assumes is the original input-string, but really is output from other work by itself or the other machine.

Assume that $L_1$ and $L_2$ are decidable languages. That is, there exists turing-machines $M_1$ and $M_2$ that respectively decides $L_1$ and $L_2$. Let us now show that the closure property holds under the different operations. In general, we will use non-deterministic TM's in our constructions, but since every non-deterministic TM has an equivalent deterministic TM, this does not make a difference for the decidability of the languages.

\textbf{b)} We need to contruct a TM M that decides the concatenation $L_1L_2$. Let M be the TM given by the following high-level description:

On input-string w: \begin{enumerate}
\item Use non-determinism to split up w into xy in all possible ways.
\item For each split up of w into xy:
\begin{enumerate}
\item Run $M_1$ on x.
\item Run $M_2$ on y.
\item Accept, if both $M_1$ and $M_2$ accepts. Otherwise, reject.  
\end{enumerate}
\end{enumerate}

We now have to argue that M accepts if $w$ is $L_1L_2$, and rejects if $w$ is not in $L_1L_2$. Since every input-string w has finite length, there are only finitely many ways to split it up into two substrings. Therefore, we have only finitely many runs of line 2. For each run, the substring $x$ is accepted by $M_1$, if $x$ is in $L_1$, and rejected if $x$ is not in $L_1$. The same goes for $y$ with respect to $M_2$. Thus, if $w\in L_1L_2$, then there exist a way to split up $w$ in $xy$ such that $x\in L_1$, and $y\in L_2$, and then there will be some computation path in M that accepts $w$. Otherwise, if $w\notin L_2$, then for all computation paths in M, $M_1$ or $M_2$ will reject and therefore M will reject. Thus, M decides $L_1L_2$.

\textbf{c)} We need to contruct a TM M that decides $(L_1)^*$. Let M be the TM given by the following high-level description.

On input-string w:
\begin{enumerate}
\item Use non-determinism to split up w into $x_1x_2...x_n$ in all possible ways, where $|x_1|=|x_2|=...=|x_n|$, and $1\leq n \leq |w|$.
\item For each split up of w into $x_1x_2...x_n$:
\begin{enumerate}
\item Run $M_1$ on $x_i$ for all $1\leq i \leq n$.
\item Accept, if all runs of $M_1$ accepts. Otherwise, reject.
\end{enumerate} 
\end{enumerate}
We now have to argue that M accepts if $w$ is $(L_1)^*$, and rejects if $w$ is not in $(L_1)^*$. Since every input-string w has finite length, there are only finitely many ways to split up w in the way described by line 1. Therefore, we have only finitely many runs of line 2. For each run, every the substring $x_i$ is accepted by $M_1$, if $x_i$ is in $L_1$ and rejected if $x_i$ is not in $L_1$. Thus, if $w\in (L_1)^*$, then there exist a way to split up $w$ in the way described by line 1 such that $x_i \in L_1$ for all $1\leq i \leq n$, and then there will be some computation path in M that accepts $w$. Otherwise, if $w\notin L_2$, then for all computation paths in M, $M_1$ will reject some substring $x_i$. Thus, M decides $(L_1)^*$.

\textbf{d)} We need to contruct a TM M that decides $\overline{L_1}$. Let M be the TM given by the following high-level description.

On input-string w:
\begin{enumerate}
\item Run $M_1$ on w. If $M_1$ accepts, then reject. Otherwise, accept. 
\end{enumerate}
We now have to argue that M accepts if $w$ is $\overline{L_1}$, and rejects if $w$ is not in $\overline{L_1}$. If $w$ is in $\overline{L_1}$, then $M_1$ rejects, and then M accepts. If $w$ is not in $\overline{L_1}$, then $w$ is in $L_1$, and then $M_1$ accepts, and then M rejects.

\textbf{e)} Since $L_1 \cap L_2=\overline{(\overline{L_1}\cup \overline{L_2})}$, then the closure under intersection follows from the closure under union and complementation.

\section*{3.18}

\end{document}

### 2017-04-30 assignments/1/paper.tex
Finished E1.3, P1.1, P1.4 a and b.

@@
 \usepackage{tikz}
+\usepackage[]{algorithm2e}
 \usepackage{parskip} 			   % no paragraph indentation
@@
 
-\title{Assignment 2}
+\title{Randomized Algorithms \\
+       Assignment 1}
 
-\author{Viktor Hansen, Asger Andersen \\ 
-Department of Computer Science \\
-University of Copenhagen}
+\author{Viktor Hansen, Simon Schleicher}
 
@@
 \begin{abstract}
-  This is the second weekly assignment for the Computability and Complexity course offered at The Department of Computer Science, Uni. Copenhagen.
+  This is the first weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
 \end{abstract}
@@
 
-\section*{2.6b}
-The complement of the language is $\left\{ w \; | \; w \in \left\{ a, b \right\}^* \right\} \setminus \left\{ a^nb^n \; | \; n \geq 0 \right\}$.
-Hence strings in the language $\left\{ \; a^nb^m \; | \; n,m \geq 0, n \neq m \right\} \subset \left\{ a^nb^m \; | \; n,m \geq 0 \right\}$ are in the complement while strings in $\left\{ \; a^nb^n \; | \; n \geq 0 \right\} \subset \left\{ a^nb^m \; | \; n,m \geq 0 \right\}$ are not, which needs to be reflected in the grammar:
-\begin{equation*}
-	\begin{split}
-S &\longrightarrow aSb \; | \; aA \; | \; bB\\
-A &\longrightarrow aA \; | \; \varepsilon \; | \; baD
-	\end{split}
-\quad\quad
-	\begin{split}
-B &\longrightarrow Bb \; | \; \varepsilon \; | \; aD\\
-D &\longrightarrow aD \; | \; bD \; | \; \varepsilon
-	\end{split}
-\end{equation*}
-We observe that at any given point a $b$ is followed by an $a$ in a derivation, the string remains in the complement, but not in $\left\{ a^nb^m \; | \; n,m \geq 0 \right\}$, regardless of any further derivation.
-This is captured in productions $A \rightarrow baD$ and $B \rightarrow aD$.
-The productions for $S$ enforce the requirement that no string in $\left\{ \; a^nb^n \; | \; n \geq 0 \right\}$ are in the complement.
-This is done by 'pumping' a string of the form $a \hdots a S b \hdots b$ and only allowing $S$ to become either $a$s, $b$s or a mix of either in which it is guaranteed that a $b$ is followed by an $a$ (productions $A \rightarrow baD$ and $B \rightarrow aD$).
-The production for $D$ derives strings in $\left\{a,b\right\}^{*}$, which is any string consisting solely of $a$s and $b$s.
+\section*{Exercise 1.2}
+Consider the graph $G=(E,V)$ consisting of two $(\left|V\right|/2)$-cliques, $L$ and $R$, connected by a single edge.
+The min-cut of $G$ has size 1.
+If at any point, the algorithm coalesces any pair of vertices that are not both from $L$ or $R$, respectively, the size of the min-cut increases in the resulting graph, and it can not be found in subsequent iterations.
 
-\section*{2.6d}
-The following grammar recognizes the language:
-\begin{equation*}
-	\begin{split}
-F &\longrightarrow LPR
-	\end{split}
-\quad\quad
-	\begin{split}
-L &\longrightarrow LQ\# \; | \; \varepsilon \\
-R &\longrightarrow \#QR \; | \; \varepsilon\\
-Q &\longrightarrow aQ \; | \; bQ \; | \; \varepsilon
-	\end{split}
-\quad\quad
-	\begin{split}
-P &\longrightarrow aPa \; | \; bPb \; | \; P'\\
-P' &\longrightarrow \#S\# \; | \; \# \; | \; a \; | \; b \; | \; \varepsilon\\
-S &\longrightarrow Q \; | \; Q \# S \\
-	\end{split}
-\end{equation*}
+\section*{Exercise 1.3}
+To obtain a Las Vegas algorithm from $A$, we run $A$ on $\Pi$ and verify the solution in time $T(n) + t(n)$.
+We repeat this process until the correct answer is given by $A$.
+Let $N$ be the random variable denoting the number of times $A$ must be run on $\Pi$ to obtain the correct answer.
+Then the expected running time of the Las Vegas algorithm is given by $\mathbf{E}\left[ \left(T(n) + t(n)\right) \cdot N \right] = \left(T(n) + t(n)\right) \cdot \mathbf{E} \left[ N \right] = \left(T(n) + t(n)\right) \cdot 1/\gamma{(n)}$ since $N$ follows a geometric distribution with parameter $\gamma{(n)}$.
 
-The production for $Q$ recognizes the language $\left\{ a, b \right\}^*$, i.e.
-each substring $x_i$.
-The language requires that for some $i,j$, $x_i=x_j^{\mathcal{R}}$.
-In case that $i=j$, $x_i=x_j=x_i^{\mathcal{R}}$ is a palindrome, which is captured by the productions for $P$ and $P' \rightarrow a \; | \; b \; | \; \varepsilon$.
-When $x_{j}=x_{i+l}^{\mathcal{R}}$ for some integer $l \neq 0$, a string in the language $s = \left\{ \; \# x_1 \# x_2 \# \hdots \# x_k \# \; | \; k \geq 1,\text{ each  } x_i \in \left\{ a, b \right\}^* \; \right\}$ exists between $x_i$ and $x_j$, i.e.
-$x_{i} s x_{j}$, which is captured by the production $P' \rightarrow \#S\#$ and those for $S$.
-The productions for $L$ (and accordingly for $R$) optionally 'pad' the language with strings in the language $\left\{ \; x_1 \# x_2 \# \hdots \# x_k \# \; | \; k \geq 0,\text{ each  } x_i \in \left\{ a, b \right\}^* \; \right\}$ (and accordingly for $R$), allowing for $i$ and $j$ to be different from $1$ and $k$.
-The start production $F \rightarrow LPR$ yields the optional left 'padding', $L$, followed by the palindrome/reversed string part, followed by optional right 'padding'.
+\section*{Problem 1.1}
+\paragraph{a)} For $k \geq 1$, let $s \in \left\{ H,T \right\}^k$ denote a $k$-coin toss sequence, s.t., for instance, $HT$ is a 2-coin toss sequence in which the first toss is a heads and the second is a tails.
+Then $HT$ and $TH$ can be viewed as the events heads and tails for a fair coin, respectively, since they are equiprobable, i.e.
+$\textbf{Pr}\left[ HT \right] = \textbf{Pr}\left[ TH \right] = p(1-p)=(1-p)p$.
+Let $N$ be the random variable denoting the number of independent 2-coin sequences needed to get either $HT$ or $TH$.
+Then $N$ follows a geometric distribution with parameter $p(1-p) + (1-p)p = 2p(1-p)$ since the events are independent, and $\mathbf{E}\left[ N \right] = 1/2p(1-p)$ which is no more than $1/p(1-p)$.
 
-\section*{2.14}
-First, the start symbol $S$ and its production is added to the grammar:
-\begin{equation*}
-	\begin{split}
-		A &\rightarrow BAB \; | \; B \; | \; \varepsilon\\
-		B &\rightarrow 00 \; | \; \varepsilon
-	\end{split}
-\quad\longrightarrow\quad
-	\begin{split}
-		\mathbf{S} &\rightarrow \mathbf{A}\\
-    	A &\rightarrow BAB \; | \; B \; | \; \varepsilon\\
-		B &\rightarrow 00 \; | \; \varepsilon
-    \end{split}
-\end{equation*}
-Next, the $B \rightarrow \varepsilon$ and $A \rightarrow \varepsilon$ productions are eliminated, yielding:
-\begin{equation*}
-	\begin{split}
-		S &\rightarrow A\\
-    	A &\rightarrow BAB \; | \; B \; | \; \bm{\varepsilon}\\
-		B &\rightarrow 00 \; | \; \bm{\varepsilon}
-    \end{split}
-\quad\longrightarrow\quad
-	\begin{split}
-		S &\rightarrow A \; | \; \bm{\varepsilon}\\
-    	A &\rightarrow BAB \; | \; B \; | \; \mathbf{AB} \; | \; \mathbf{BA} \; | \; \mathbf{A} \; | \; \mathbf{BB}\\
-		B &\rightarrow 00
-    \end{split}
-\end{equation*}
-Now, unit rules $S \rightarrow A$, $A \rightarrow B$ and $A \rightarrow A$ are removed:
-\begin{equation*}
-	\begin{split}
-		S &\rightarrow \mathbf{A} \; | \; \varepsilon\\
-    	A &\rightarrow BAB \; | \; \mathbf{B} \; | \; AB \; | \; BA \; | \; \mathbf{A} \; | \; BB\\
-		B &\rightarrow 00
-    \end{split}
-\quad\longrightarrow\quad
-	\begin{split}
-		S &\rightarrow \mathbf{BAB \; | \; 00 \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon}\\
-    	A &\rightarrow BAB \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB\\
-		B &\rightarrow 00
-    \end{split}
-\end{equation*}
-Finally, the remaining rules are converted into Chomsky normal form by adding additional variables and rules to the grammar, yielding:
-\begin{equation*}
-	\begin{split}
-		S &\rightarrow \mathbf{BA}B \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon\\
-    	A &\rightarrow \mathbf{BA}B \; | \; \mathbf{00} \; | \; AB \; | \; BA \; | \; BB\\
-		B &\rightarrow 00
-    \end{split}
-\quad\longrightarrow\quad
-	\begin{split}
-		S &\rightarrow \mathbf{C}B \; | \; \mathbf{DD} \; | \; AB \; | \; BA \; | \; BB \; | \; \varepsilon\\
-    	A &\rightarrow \mathbf{C}B \; | \; \mathbf{DD} \; | \; AB \; | \; BA \; | \; BB\\
-		\mathbf{C} &\rightarrow \mathbf{BA}\\
-		\mathbf{B} &\rightarrow \mathbf{DD}\\
-		\mathbf{D} &\rightarrow \mathbf{0}
-    \end{split}
-\end{equation*}
+\paragraph{b)} Assume (without loss of generality) that $k$ is even.
+Given a sequence $s \in \left\{ H,T \right\}^k$, an extension of the above scheme does the following: 1) it scans through $s$ two elements at a time and extracts tosses in the manner described in a).
+2) it subsequently removes the pairs corresponding to those tosses from the $s$.
+It is easily verified that the resulting list has the form $\left\{ HH,TT \right\}^{k'}$ where $k' \leq k$.
+3) it removes every other member of $s$, once again producing a sequence of the form $\left\{ H,T \right\}^k$.
+Steps 1), 2) and 3) are repeated until the list is empty, and the extracted tosses are reported.
 
-\section*{2.42}
-\textbf{a)} Assume that the language $L_1 = \{0^n1^n0^n1^n| n\geq 0 \}$ is context-free.
-With this assumption, we get from the pumping lemma for context-free languages to that there exists a natural number p, such that all strings in $L_1$, which are at least p long, satisfy the three conditions in the lemma.
-But consider the string $s=0^p1^p0^p1^p$.
-Since $s$ is in $L_1$ and is longer than p, we know that $s$ satisfies the three conditions of the lemma.
-That is, we can write $s=uvxyz$ such that $|vy|>0$ and $|vxy|\leq p$ and $uv^ixy^iz$ is in $L_1$ for all $i\geq 0$.
-But since $|vxy|\leq p$, we have from the definition of $s$ that either $v$ and $y$ do not contain any 0's or they do not contain any 1's.
-Since $|vy|>0$, we have that at least one of $v$ or $y$ contains at least one 0 or 1.
-For that reason, when we pump $s$ to $uv^2xy^2z$ either the number of 0's will grow, while the number of 1's will not change, or the number of 1's will grow, while the number of 0's will not change.
-In both cases, the pumped string will have a different number of 0's and 1's.
-But all the strings in $L_1$ have the same number of 0's and 1's.
-Therefore, the pumped string is not in $L_1$.
-But from the pumping lemma, we know that $uv^2xy^2z$ is in $L_1$.
-This is a contradiction, which we have derived from our assumption that $L_1$ is context-free.
-This assumption is therefore false.
-$L_1$ is therefore not context-free.
+\section*{Problem 1.4}
+\paragraph{a)} One solution generates a list of all $n!$ possible permutations and subsequently generates a random index into the list.
+Generating one such index requires $\lceil \log_2(n!) \rceil$ random bits and can be generated s.t.
+it falls within bounds of the list with probability $n!
+/ 2^{\lceil \log_2(n!) \rceil} = \log_2(n!) / \lceil \log_2(n!) \rceil \geq \log_2(n!) / (\log_2(n!)+1) \geq 1/2$.
+Let $N$ be the random variable denoting the number of indices that need to be generated to make the index falls within the bounds of the list.
+Then the expected total number of bits needed to generate such an index is given by $\textbf{E}\left[ N \cdot \lceil \log_2(n!) \rceil \right] = \textbf{E}\left[ N \right] \cdot \lceil \log_2(n!) \rceil \leq 2 \cdot \lceil \log_2(n!) \rceil \leq 2\log_2(n!) + 1 \leq (n+1) \log_2 \left( \frac{n+1}{e} \right) + 2 = O(n \log n)$ since $N$ is geometrically distributed with parameter $\geq 1/2$.
+For the lower bound, we get $\textbf{E}\left[ N \right] \cdot \lceil \log_2(n!) \rceil \geq 1 \cdot \lceil \log_2(n!) \rceil = \Theta(n \log_2 n)$ by a similar reasoning.
+The time and space complexities are an impractical $\Theta(n!)$.
 
-\textbf{b)} Assume that the language $L_2 = \{t_1\#...\#t_k| k\geq 2,\  t_i\in \{ a,b \}^*,\ t_i = t_j \text{ for some } i\neq j \}$ is context-free.
-With this assumptions, we now get from the pumping lemma for context-free languages to that there exists a natural number p, such that all strings in $L_2$, which are at least p long, satisfy the three conditions in the lemma.
-But consider the string $s=a^pb^p \# a^pb^p$.
-Since $s$ is in $L_2$ and is longer than p, we know that $s$ satisfies the three conditions of the lemma.
-That is, we can write $s=uvxyz$ such that $|vy|>0$ and $|vxy|\leq p$ and $uv^ixy^iz$ is in $L_2$ for all $i\geq 0$.
+Another approach creates a bijection $f : [1 \hdots n] \rightarrow [1 \hdots n]$ by randomly assigning $\lceil \log_2 n \rceil$ bits to each element in the domain of $f$ corresponding to the elements of the range.
+Let $\varepsilon_i$ denote the event that the $i$th element of the domain is assigned a bitstring corresponding to a valid element in the range i.e.
+s.t.
+$f$ is a bijection.
+There are $n-i+1$ valid indices and thus
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon_i \right] = \frac{n-i+1}{2^{\lceil \log_2 n \rceil}} \geq \frac{1}{2^{\lceil \log_2 n \rceil}} \geq \frac{1}{2^{\log_2 (n) + 1}} = \frac{1}{2n}
+\end{align*}
+Let $N_i$ be the random variable denoting the number of bitstrings required to generate a valid element of the range for the $i$th element in the domain.
+Then, the total expected number of bits required is
+\begin{align*}
+\mathbf{E} \left[ \sum_{i=1}^{n} N_i \cdot \lceil \log_2 n \rceil \right] = \sum_{i=1}^{n} \mathbf{E}\left[ N_i \right] \cdot \lceil \log_2 n \rceil \leq \sum_{i=1}^{n} 2n \cdot \lceil \log_2 n \rceil = O(n^2 \log_2 n)
+\end{align*}
+For the lower bound, we note that we need at least one try for each element in the domain and get
+\begin{align*}
+\sum_{i=1}^{n} \mathbf{E}\left[ N_i \right] \cdot \lceil \log_2 n \rceil \geq \sum_{i=1}^{n} 1 \cdot \lceil \log_2 n \rceil = \Omega(n \log_2 n)
+\end{align*}
+For the running time, the algorithm checks whether a generated element is already used in the permutation in $\Theta(n)$ time, yielding expected running times of $\Omega(n^2)$ and $O(n^3)$.
 
-All the ways of splitting up $s$ as $uvxyz$ can be categorized with these cases: 
-\begin{enumerate}
-\item None of $v$ or $y$ contains the $\#$ symbol:
-\begin{enumerate}
-\item $v$ and $y$ both contain symbols only from the same side of $\#$.
-\item $v$ contains symbols only from the left side of $\#$, and $y$ contains symbols only from the right side of $\#$.
-\end{enumerate} 
-\item One of $v$ or $y$ contains the $\#$ symbol.
-\end{enumerate}
+\paragraph{b)} The claim is correct, since there is a 0 probability of a collisions occurring if the numbers are picked from the interval, uniformly, at random.
+However an unbounded number of bits are required to represent such a number, and hence a lazy approach, such as the one described in c) is required.
 
-Let us now argue why for each case we cannot find $v$ and $y$ such that $s=uvxyz$ can be pumped. \begin{enumerate}
-\item
-\begin{enumerate}
-\item Since none of $v$ or $y$ contain the $\#$, then $uv^2xy^2z$ still only contains one $\#$.
-Since $v$ and $y$ cannot both be empty strings and they both only contains symbols from one side of the $\#$, then the pumping only effects one side of the $\#$.
-Therefore $uv^2xy^2z$ will have the form $t_1\#t_2$, where $t_1\neq t_2$.
-This string is not in $L_2$.
-\item Since none of $v$ or $y$ contain the $\#$, then $uv^2xy^2z$ still only contains one $\#$.
-Since $v$ and $y$ contain symbols from different sides of $\#$ and none of the contains $\#$, then $x$ must contain $\#$.
-Since $|vxy|\leq p$, then we can conclude that $v$ cannot contain other symbols than $b's$, and $y$ cannot contain other symbols than $a's$.
-Since $v$ and $y$ cannot both be empty strings, then we can now conclude that $uv^2xy^2z$ will have the form $t_1\#t_2$, where $t_1\neq t_2$.
-This string is not in $L_2$.
-\end{enumerate}
-\item Since $v$ or $y$ contains the $\#$, then $uv^0xy^0z$ will not have any $\#$.
-In other words, it will just have the form $t_1$, where $t_1$ is some string in $\{a,b\}^*$.
-This string is not in $L_2$.
-\end{enumerate}
+\paragraph{c)}
 
-\section*{2.54}
-We consider the possible ways to divide the string $s=1^0\#1^1\#1^2\# \hdots \#1^p$ according to the pumping lemma for CFGs.
-Obviously $s$ is in the language and the pumping lemma states that since $\left| s \right| \geq p$, it can be divided it into parts $uvxyz$, st.
-$\left| vxy \right| \leq p$, $\left| vy \right| \geq 0$ and $uv^ixy^iz$ is in the language for $i \geq 0$.
-
-Choosing either $v$ or $y$ as $\#$, pumping it up would yield more than one empty sequence between the sharps in $s$, and the resulting string would not be in the language.
-Choosing either $v$ or $u$ as any substring containing at least one sharp would also lead to a contradiction when pumping up (or down), by similar reasoning.
-Choosing either $v$ or $u$ as all ones will lead to a contradiction when pumping down, since this will result in a sequence of ones that already exists elsewhere.
-Specifically, choosing a length $k \leq p$ all-ones substring of any all-ones substring and pumping down will yield another all-ones substring of size at most $p-k$.
-But all such strings $1^{p-k}$ for $k \leq p$ are already a substring between sharps in $s$, leading to the contradiction and hence the grammar is not context free.
-
-%\begin{figure}[!hbt]
-%\center
-%\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,semithick]
-%  \tikzstyle{every state}=[fill=red,draw=none,text=white]
-
-%  \node[initial,state] (A)                    {$q_a$};
-%  \node[state]         (B) [above right of=A] {$q_b$};
-%  \node[state]         (D) [below right of=A] {$q_d$};
-%  \node[state]         (C) [below right of=B] {$q_c$};
-%  \node[state]         (E) [below of=D]       {$q_e$};
-
-%  \path (A) edge              node {0,1,L} (B)
-%            edge              node {1,1,R} (C)
-%        (B) edge [loop above] node {1,1,L} (B)
-%            edge              node {0,1,L} (C)
-%        (C) edge              node {0,1,L} (D)
-%            edge [bend left]  node {1,0,R} (E)
-%        (D) edge [loop below] node {1,1,R} (D)
-%            edge              node {0,1,R} (A)
-%        (E) edge [bend left]  node {1,0,R} (A);
-%\end{tikzpicture}
-%  \caption{Output of running the program, showing the values produced by each %step.\label{lst:output}}
-%\end{figure}
-
-\section*{3.16}
-The book's solution to 3.16a) is a quite high-level description of how to combine a TM $M_1$ deciding a language $L_1$ and a TM $M_2$ deciding the language $L_2$ into a new TM $M$ that decides the union of the two languages.
-Our solution will follow this high-level style (that is, not specifying the details of the 7-tuples that describes the different turing-machines).
-We will be more careful than the book in the sense that we use non-determinism to make it a little more explicit how the new machine M makes sure that one of the machines $M_i$ does not end up working on something that it assumes is the original input-string, but really is output from other work by itself or the other machine.
-
-Assume that $L_1$ and $L_2$ are decidable languages.
-That is, there exists turing-machines $M_1$ and $M_2$ that respectively decides $L_1$ and $L_2$.
-Let us now show that the closure property holds under the different operations.
-In general, we will use non-deterministic TM's in our constructions, but since every non-deterministic TM has an equivalent deterministic TM, this does not make a difference for the decidability of the languages.
-
-\textbf{b)} We need to contruct a TM M that decides the concatenation $L_1L_2$. Let M be the TM given by the following high-level description:
-
-On input-string w: \begin{enumerate}
-\item Use non-determinism to split up w into xy in all possible ways.
-\item For each split up of w into xy:
-\begin{enumerate}
-\item Run $M_1$ on x.
-\item Run $M_2$ on y.
-\item Accept, if both $M_1$ and $M_2$ accepts. Otherwise, reject.  
-\end{enumerate}
-\end{enumerate}
-
-We now have to argue that M accepts if $w$ is $L_1L_2$, and rejects if $w$ is not in $L_1L_2$.
-Since every input-string w has finite length, there are only finitely many ways to split it up into two substrings.
-Therefore, we have only finitely many runs of line 2.
-For each run, the substring $x$ is accepted by $M_1$, if $x$ is in $L_1$, and rejected if $x$ is not in $L_1$.
-The same goes for $y$ with respect to $M_2$.
-Thus, if $w\in L_1L_2$, then there exist a way to split up $w$ in $xy$ such that $x\in L_1$, and $y\in L_2$, and then there will be some computation path in M that accepts $w$.
-Otherwise, if $w\notin L_2$, then for all computation paths in M, $M_1$ or $M_2$ will reject and therefore M will reject.
-Thus, M decides $L_1L_2$.
-
-\textbf{c)} We need to contruct a TM M that decides $(L_1)^*$. Let M be the TM given by the following high-level description.
-
-On input-string w:
-\begin{enumerate}
-\item Use non-determinism to split up w into $x_1x_2...x_n$ in all possible ways, where $|x_1|=|x_2|=...=|x_n|$, and $1\leq n \leq |w|$.
-\item For each split up of w into $x_1x_2...x_n$:
-\begin{enumerate}
-\item Run $M_1$ on $x_i$ for all $1\leq i \leq n$.
-\item Accept, if all runs of $M_1$ accepts. Otherwise, reject.
-\end{enumerate} 
-\end{enumerate}
-We now have to argue that M accepts if $w$ is $(L_1)^*$, and rejects if $w$ is not in $(L_1)^*$.
-Since every input-string w has finite length, there are only finitely many ways to split up w in the way described by line 1.
-Therefore, we have only finitely many runs of line 2.
-For each run, every the substring $x_i$ is accepted by $M_1$, if $x_i$ is in $L_1$ and rejected if $x_i$ is not in $L_1$.
-Thus, if $w\in (L_1)^*$, then there exist a way to split up $w$ in the way described by line 1 such that $x_i \in L_1$ for all $1\leq i \leq n$, and then there will be some computation path in M that accepts $w$.
-Otherwise, if $w\notin L_2$, then for all computation paths in M, $M_1$ will reject some substring $x_i$.
-Thus, M decides $(L_1)^*$.
-
-\textbf{d)} We need to contruct a TM M that decides $\overline{L_1}$. Let M be the TM given by the following high-level description.
-
-On input-string w:
-\begin{enumerate}
-\item Run $M_1$ on w. If $M_1$ accepts, then reject. Otherwise, accept. 
-\end{enumerate}
-We now have to argue that M accepts if $w$ is $\overline{L_1}$, and rejects if $w$ is not in $\overline{L_1}$.
-If $w$ is in $\overline{L_1}$, then $M_1$ rejects, and then M accepts.
-If $w$ is not in $\overline{L_1}$, then $w$ is in $L_1$, and then $M_1$ accepts, and then M rejects.
-
-\textbf{e)} Since $L_1 \cap L_2=\overline{(\overline{L_1}\cup \overline{L_2})}$, then the closure under intersection follows from the closure under union and complementation.
-
-\section*{3.18}
+\section*{Problem 1.8}
 

### 2017-04-30 assignments/1/paper.tex
Started typing in 1.2

@@
 \section*{Exercise 1.2}
-Consider the graph $G=(E,V)$ consisting of two $(\left|V\right|/2)$-cliques, $L$ and $R$, connected by a single edge.
+Consider the graph $G=(V,E)$ consisting of two $(\left|V\right|/2)$-cliques, $L$ and $R$, connected by a single edge (assume that $|V|$ is even, without loss of generality).
 The min-cut of $G$ has size 1.
 If at any point, the algorithm coalesces any pair of vertices that are not both from $L$ or $R$, respectively, the size of the min-cut increases in the resulting graph, and it can not be found in subsequent iterations.
+Let $\varepsilon$ denote the event that a pair of vertices from either clique are chosen.
+Then
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon \right] = \frac{\binom{|L|}{2} + \binom{|R|}{2}}{\binom{|L|+|R|}{2}}
+\end{align*}
 
@@
 
-\paragraph{b)} The claim is correct, since there is a 0 probability of a collisions occurring if the numbers are picked from the interval, uniformly, at random.
+\paragraph{b)} The claim is correct, since there is a 0 probability of collisions occurring if the numbers are picked from the interval, uniformly, at random.
 However an unbounded number of bits are required to represent such a number, and hence a lazy approach, such as the one described in c) is required.

### 2017-05-07 assignments/1/paper.tex
Update paper.tex
Added assignment 2 stub.

@@
 
-\author{Viktor Hansen, Simon Schleicher}
+\author{Viktor Hansen \& Simon Rueskov Schleicher}
 
@@
 If at any point, the algorithm coalesces any pair of vertices that are not both from $L$ or $R$, respectively, the size of the min-cut increases in the resulting graph, and it can not be found in subsequent iterations.
-Let $\varepsilon$ denote the event that a pair of vertices from either clique are chosen.
+Let $\varepsilon_i$ denote the event that a pair of vertices from either clique is chosen in the $i$th iteration and $L_{i}$ and $R_{i}$ denote the cliques before the $i$th iteration.
 Then
 \begin{align*}
-\mathbf{Pr}\left[ \varepsilon \right] = \frac{\binom{|L|}{2} + \binom{|R|}{2}}{\binom{|L|+|R|}{2}}
+\mathbf{Pr}\left[ \varepsilon_i \; | \, \cap_{j=1}^{i-1} \varepsilon_j \right] &= \frac{\binom{|L_i|}{2} + \binom{|R_i|}{2}}{\binom{|L_i|+|R_i|}{2}}
+= \frac{\frac{|L_i|!}{2!(|L_i|-2)!} + \frac{|R_i|!}{2!(|R_i|-2)!}}{\frac{(|L_i|+|R_i|)!}{2!((|L_i|+|R_i|)-2)!}}
+= \frac{|L_i|(|L_i|-1) + |R_i|(|R_i|-1)}{(|L_i|+|R_i|)((|L_i|+|R_i|)-1)} \\
+&= \frac{|L_i|^2+ |R_i|^2 - (|L_i| +|R_i|)}{(|L_i|+|R_i|)^2-(|L_i|+|R_i|)}
+= \frac{(|L_i| + |R_i|)^2 - 2|L_i||R_i| - (|L_i| +|R_i|)}{(|L_i|+|R_i|)^2-(|L_i|+|R_i|)} \\
+&= 1 - \frac{2|L_i||R_i|}{(|L_i|+|R_i|)^2-(|L_i|+|R_i|)} \leq 1 -  \frac{2|L_i||R_i|}{n^2}
 \end{align*}
+Now for any $i$, s.t. $1 \leq i \leq n/x$ for some $x > 2$, we have $|R_i|,|L_i| \geq n/x$ and
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon_i \; | \, \cap_{j=1}^{i-1} \varepsilon_j \right] &\leq 1 - \frac{2|L_i||R_i|}{n^2} \leq 1 - \frac{2(n/x)^2}{n^2} = 1 - \frac{2}{x^2}
+\end{align*}
+Finally we get
+\begin{align*}
+\mathbf{Pr}\left[ \cap_{i=1}^{n} \varepsilon_i \right] \leq \mathbf{Pr}\left[ \cap_{i=1}^{\lceil n/x \rceil} \varepsilon_i \right] \leq \left( 1 - \frac{2}{x^2} \right)^{\lceil n/x \rceil}
+\end{align*}
+which is exponentially small in $n$ for any choice of $x > 2$.
 
@@
 It is easily verified that the resulting list has the form $\left\{ HH,TT \right\}^{k'}$ where $k' \leq k$.
-3) it removes every other member of $s$, once again producing a sequence of the form $\left\{ H,T \right\}^k$.
+3) it removes every other member of $s$, once again producing a sequence of the form $\left\{ H,T \right\}^{k'}$.
 Steps 1), 2) and 3) are repeated until the list is empty, and the extracted tosses are reported.
@@
 
+\pagebreak
+
+\section*{Summary}
+
 \end{document}

### 2017-05-07 assignments/2/paper.tex
Added assignment 2 stub.

--- 2017-04-30 assignments/1/paper.tex
@@
 \title{Randomized Algorithms \\
-       Assignment 1}
+       Assignment 2}
 
-\author{Viktor Hansen, Simon Schleicher}
+\author{Viktor Hansen \& Simon Rueskov Schleicher}
 
@@
 \begin{abstract}
-  This is the first weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
+  This is the second weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
 \end{abstract}
@@
 
-\section*{Exercise 1.2}
-Consider the graph $G=(E,V)$ consisting of two $(\left|V\right|/2)$-cliques, $L$ and $R$, connected by a single edge.
-The min-cut of $G$ has size 1.
-If at any point, the algorithm coalesces any pair of vertices that are not both from $L$ or $R$, respectively, the size of the min-cut increases in the resulting graph, and it can not be found in subsequent iterations.
+\section*{Exercise ???}
 
-\section*{Exercise 1.3}
-To obtain a Las Vegas algorithm from $A$, we run $A$ on $\Pi$ and verify the solution in time $T(n) + t(n)$.
-We repeat this process until the correct answer is given by $A$.
-Let $N$ be the random variable denoting the number of times $A$ must be run on $\Pi$ to obtain the correct answer.
-Then the expected running time of the Las Vegas algorithm is given by $\mathbf{E}\left[ \left(T(n) + t(n)\right) \cdot N \right] = \left(T(n) + t(n)\right) \cdot \mathbf{E} \left[ N \right] = \left(T(n) + t(n)\right) \cdot 1/\gamma{(n)}$ since $N$ follows a geometric distribution with parameter $\gamma{(n)}$.
+\pagebreak
 
-\section*{Problem 1.1}
-\paragraph{a)} For $k \geq 1$, let $s \in \left\{ H,T \right\}^k$ denote a $k$-coin toss sequence, s.t., for instance, $HT$ is a 2-coin toss sequence in which the first toss is a heads and the second is a tails.
-Then $HT$ and $TH$ can be viewed as the events heads and tails for a fair coin, respectively, since they are equiprobable, i.e.
-$\textbf{Pr}\left[ HT \right] = \textbf{Pr}\left[ TH \right] = p(1-p)=(1-p)p$.
-Let $N$ be the random variable denoting the number of independent 2-coin sequences needed to get either $HT$ or $TH$.
-Then $N$ follows a geometric distribution with parameter $p(1-p) + (1-p)p = 2p(1-p)$ since the events are independent, and $\mathbf{E}\left[ N \right] = 1/2p(1-p)$ which is no more than $1/p(1-p)$.
-
-\paragraph{b)} Assume (without loss of generality) that $k$ is even.
-Given a sequence $s \in \left\{ H,T \right\}^k$, an extension of the above scheme does the following: 1) it scans through $s$ two elements at a time and extracts tosses in the manner described in a).
-2) it subsequently removes the pairs corresponding to those tosses from the $s$.
-It is easily verified that the resulting list has the form $\left\{ HH,TT \right\}^{k'}$ where $k' \leq k$.
-3) it removes every other member of $s$, once again producing a sequence of the form $\left\{ H,T \right\}^k$.
-Steps 1), 2) and 3) are repeated until the list is empty, and the extracted tosses are reported.
-
-\section*{Problem 1.4}
-\paragraph{a)} One solution generates a list of all $n!$ possible permutations and subsequently generates a random index into the list.
-Generating one such index requires $\lceil \log_2(n!) \rceil$ random bits and can be generated s.t.
-it falls within bounds of the list with probability $n!
-/ 2^{\lceil \log_2(n!) \rceil} = \log_2(n!) / \lceil \log_2(n!) \rceil \geq \log_2(n!) / (\log_2(n!)+1) \geq 1/2$.
-Let $N$ be the random variable denoting the number of indices that need to be generated to make the index falls within the bounds of the list.
-Then the expected total number of bits needed to generate such an index is given by $\textbf{E}\left[ N \cdot \lceil \log_2(n!) \rceil \right] = \textbf{E}\left[ N \right] \cdot \lceil \log_2(n!) \rceil \leq 2 \cdot \lceil \log_2(n!) \rceil \leq 2\log_2(n!) + 1 \leq (n+1) \log_2 \left( \frac{n+1}{e} \right) + 2 = O(n \log n)$ since $N$ is geometrically distributed with parameter $\geq 1/2$.
-For the lower bound, we get $\textbf{E}\left[ N \right] \cdot \lceil \log_2(n!) \rceil \geq 1 \cdot \lceil \log_2(n!) \rceil = \Theta(n \log_2 n)$ by a similar reasoning.
-The time and space complexities are an impractical $\Theta(n!)$.
-
-Another approach creates a bijection $f : [1 \hdots n] \rightarrow [1 \hdots n]$ by randomly assigning $\lceil \log_2 n \rceil$ bits to each element in the domain of $f$ corresponding to the elements of the range.
-Let $\varepsilon_i$ denote the event that the $i$th element of the domain is assigned a bitstring corresponding to a valid element in the range i.e.
-s.t.
-$f$ is a bijection.
-There are $n-i+1$ valid indices and thus
-\begin{align*}
-\mathbf{Pr}\left[ \varepsilon_i \right] = \frac{n-i+1}{2^{\lceil \log_2 n \rceil}} \geq \frac{1}{2^{\lceil \log_2 n \rceil}} \geq \frac{1}{2^{\log_2 (n) + 1}} = \frac{1}{2n}
-\end{align*}
-Let $N_i$ be the random variable denoting the number of bitstrings required to generate a valid element of the range for the $i$th element in the domain.
-Then, the total expected number of bits required is
-\begin{align*}
-\mathbf{E} \left[ \sum_{i=1}^{n} N_i \cdot \lceil \log_2 n \rceil \right] = \sum_{i=1}^{n} \mathbf{E}\left[ N_i \right] \cdot \lceil \log_2 n \rceil \leq \sum_{i=1}^{n} 2n \cdot \lceil \log_2 n \rceil = O(n^2 \log_2 n)
-\end{align*}
-For the lower bound, we note that we need at least one try for each element in the domain and get
-\begin{align*}
-\sum_{i=1}^{n} \mathbf{E}\left[ N_i \right] \cdot \lceil \log_2 n \rceil \geq \sum_{i=1}^{n} 1 \cdot \lceil \log_2 n \rceil = \Omega(n \log_2 n)
-\end{align*}
-For the running time, the algorithm checks whether a generated element is already used in the permutation in $\Theta(n)$ time, yielding expected running times of $\Omega(n^2)$ and $O(n^3)$.
-
-\paragraph{b)} The claim is correct, since there is a 0 probability of a collisions occurring if the numbers are picked from the interval, uniformly, at random.
-However an unbounded number of bits are required to represent such a number, and hence a lazy approach, such as the one described in c) is required.
-
-\paragraph{c)}
-
-\section*{Problem 1.8}
+\section*{Summary}
 

### 2017-05-07 assignments/2/paper.tex
Update paper.tex

@@
 
+\subsection*{3.1}
+We define $ee(m)$ to be the expected number of empty bins when m balls are thrown into n bins. We claim that $$ee(m)=\frac{(n-1)^m}{n^{m-1}}$$ and prove this by induction on the number m:\\
+\\
+First the base case. When 0 balls are thrown into n bins we of course expect to have n empty bins. We have $$ee(0)=\frac{(n-1)^0}{n^{0-1}}=\frac{1}{n^{-1}}=n$$ so the base case holds.\\
+\\
+Now for the induction step we assume that $ee(m)=\frac{(n-1)^m}{n^{m-1}}$ and need to prove $ee(m)=\frac{(n-1)^{m+1}}{n^m}$.
+Since the number of empty bins after one more ball can only be the same or one less than before that ball, we get that the expected number of empty bins after $m+1$ balls are equal to the expected number of empty bins after $m$ balls minus the probability that we get 1 less empty bin.
+So we get \begin{align*}ee(m+1)&=ee(m)-\frac{ee(m)}{n}=\frac{(n-1)^m}{n^{m-1}}-\frac{\left(\frac{(n-1)^m}{n^{m-1}}\right)}{n}\\&=\frac{n(n-1)^m}{n^m}-\frac{(n-1)^m}{n^m}=\frac{(n-1)(n-1)^m}{n^m}=\frac{(n-1)^{m+1}}{n^m}\end{align*}
+And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.
+
 \pagebreak

### 2017-05-07 assignments/2/paper.tex
Update paper.tex

@@
 So we get \begin{align*}ee(m+1)&=ee(m)-\frac{ee(m)}{n}=\frac{(n-1)^m}{n^{m-1}}-\frac{\left(\frac{(n-1)^m}{n^{m-1}}\right)}{n}\\&=\frac{n(n-1)^m}{n^m}-\frac{(n-1)^m}{n^m}=\frac{(n-1)(n-1)^m}{n^m}=\frac{(n-1)^{m+1}}{n^m}\end{align*}
-And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.
+And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.\\
+\\
+For $m=n$ we get $$ee(n)=\frac{(n-1)^n}{n^{n-1}}=n\frac{(n-1)^n}{n^n}=n\left(\frac{n-1}{n}\right)^n=n\left(n-\frac{1}{n}\right)^n$$ And since $\left(n-\frac{1}{n}\right)^n$ goes towards $e$ when n goes towards infinity, we get that for large n the expected number of empty bins approaches $\frac{n}{e}$ as we wanted to show.
 

### 2017-05-08 assignments/2/paper.tex
Added summary.

--- 2017-05-07 assignments/2/paper.tex
@@
 \newtheorem{lem}[thm]{Lemma}
+\newtheorem{dfn}[thm]{Definition}
 \newtheorem{prop}[thm]{Proposition}
@@
 
-\section*{Summary}
+\section{Summary}
+\subsection{Balls into bins}
+Given $n$ balls and $n$ bins, and ball $i$ goes into bin $j$ with probability $1/n$.
+Then the expected number of balls in bin $j$ is 1.
+Let $\varepsilon^=_j(k)$ denote the event where bin $j$ receives exactly $k$ balls.
+Then
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon^=_j(k) \right] = \binom{n}{k}\frac{1}{n}^n \left( 1-\frac{1}{n} \right)^{n-k} \leq \binom{n}{k}\frac{1}{n}^n \leq \binom{ne}{k}^k \left( \frac{1}{n} \right)^n = \left(\frac{e}{k}\right)^k
+\end{align*}
+Now, let $\varepsilon^\geq_j(k)$ denote the event where bin $j$ receives $k$ or more balls, then
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon^\geq_j(k) \right] = \sum_{i=k}^{n} \mathbf{Pr}\left[ \varepsilon^=_j(i) \right] \leq \sum_{i=k}^{n} \left(\frac{e}{i}\right)^i \leq \sum_{i=k}^{n} \left(\frac{e}{k}\right)^i \leq \left(\frac{e}{k}\right)^k \sum_{i=0}^{\infty} \left( \frac{e}{k} \right)^i = \left( \frac{e}{k} \right)^k \frac{1}{1-\frac{e}{k}}
+\end{align*}
+Choosing $k^*=\frac{3 \ln n}{\ln \ln n}$ and inserting it into the above yields
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon_j^\geq(k^*) \right] \leq \frac{1}{n^2}.
+\end{align*}
+The probability of any bin receiving more than $k^*$ balls is given by
+\begin{align*}
+\mathbf{Pr}\left[ \cup_{j=1}^{n} \varepsilon_j^\geq(k^*) \right] \leq \sum_{j=1}^n \mathbf{Pr}\left[ \varepsilon_j^\geq(k^*) \right] \leq n \frac{1}{n^2} = \frac{1}{n}
+\end{align*}
+
+\subsection{Markov's Inequality}
+Given a non-negative random variable $Y$ and $t > 0$, then
+\begin{align*}
+\mathbf{Pr}\left[ Y \geq t \mathbf{E}\left[ Y \right] \right] \leq \frac{1}{t} \;\; \mathrm{ and } \;\; \mathbf{Pr}\left[ Y \geq t \right] \leq \frac{\mathbf{E}\left[ Y \right]}{t}
+\end{align*}
+To show the rightmost form, we get
+\begin{align*}
+\mathbf{E}\left[ Y \right] = \sum_{i} i \cdot \mathbf{Pr}\left[ Y=i \right] \geq \sum_{i \geq t} i \cdot \mathbf{Pr}\left[ Y=i \right] \geq t \cdot \sum_{i \geq t} \mathbf{Pr}\left[ Y=i \right] \geq t \cdot \mathbf{Pr}\left[ Y \geq i \right]
+\end{align*}
+The leftmost form follows by substituting $t$ by $t \cdot \mathbf{E}\left[ Y \right]$ into the rightmost form.
+
+\subsection{Chebyshev's Inequality}
+Given rand.
+var.
+$X$, define the variance of $X$, denoted by $\sigma^2$, as $\sigma^2 = \mathbf{E}\left[ \left( X - \mu_X \right)^2 \right]$, where $\mu_X = \mathbf{E} \left[ X \right]$.
+Now, given a random variable $X$ and $t > 0$, then
+\begin{align*}
+\mathbf{Pr}\left[ \left| X-\mu_X \right| \geq t\sigma_X \right] &= \mathbf{Pr}\left[ \left( X-\mu_X \right)^2 \geq t^2\sigma_X^2 \right] \\
+&= \mathbf{Pr}\left[ \left( X-\mu_X \right)^2 \geq t^2 \mathbf{E}\left[\left( X-\mu_X \right)^2\right] \right] \\
+&= \mathbf{Pr}\left[ Y \geq t^2\mathbf{E}\left[Y\right] \right] \\
+&= \frac{1}{t^2}
+\end{align*}
+The last step follows from applying Markov's inequality.
+
+\subsection{Lazy Selection}
+Refer to the algorithm given in the book.
+Let $S(k)$ denote the $k$th smallest element in $S$ and $r_{S(a)}$ denote the rank of $a$ in $S$.
+Any known det.
+alg.
+requires $\geq 3n$ comparisons to identify $S_{k}$.
+We want to show that \textsc{LazySelect} finds $S(k)$ with $2n + O(n)$ comparisons with prob.
+$1-O(n^{-1/4})$.
+The algorithm can err.
+if $S(k) < a$.
+For this to happen, fewer than $\mathcal{l}$ samples are smaller than $S(k)$ for $i=1, \hdots, n^{3/4}$, $X_i=1$ if $i$th sample $\leq S(k)$, and $X_i = 0$ otherwise.
+Let $X=\sum^{n^{3/4}}_{i=1} X_i$.
+We will bound the error probability using Chebyshev's inequality.
+We get $\mu_X = \sum^{n^{3/4}}_{i=1} \mu_{X_i} = n^{3/4}\cdot \frac{k}{n}=kn^{-1/4}$ and $\sigma_X^2 = \sum^{n^{3/4}}_{i=1} \sigma^2_{X_i} = n^{3/4}\frac{k}{n}\left( 1-\frac{k}{n} \right) \leq \frac{1}{4}n^{3/4}$ implying $\sigma_X \leq \frac{1}{2}n^{3/8}$ and
+\begin{align*}
+\mathbf{Pr} \left[ X \leq l \right] &= \mathbf{Pr} \left[ X \leq \mu_X - \sqrt{n} \right] = \mathbf{Pr} \left[ \mu_X - X \geq \sqrt{n} \right] \leq \mathbf{Pr} \left[ \left| \mu_X - X \right| \geq \sqrt{n} \right] \\
+&\leq \mathbf{Pr} \left[ \left| \mu_X - X \right| \geq 2n^{1/8}\sigma_X \right] \leq \frac{1}{4n^{1/4}} = O(n^{-1/4})
+\end{align*}
+and hence the success probability is $\geq 1-O(n^{-1/4})$.
+
+\subsection{Two-point Sampling}
+\begin{lem}
+Let $a,b \in \mathbb{Z}_n$ where $n$ is a prime be picked independently and uniformly at random.
+Let $y_i = (a_i + b) \mod n$ for $i = 0, \hdots, n-1$.
+Then for all distinct $i,j \in \mathbb{Z}_n$, $y_i$ and $y_j$ are pairwise ind.
+and uniformly distributed in $\mathbb{Z}_n$.
+\\
+\end{lem}
+
+\begin{dfn}
+Variables $X_1, \hdots, X_n$ are pairwise ind. iff. for all distinct $i,j$ and all $x_y$, $\mathbf{Pr}\left[ X_i = x \; | \; X_j = y \right] = \mathbf{Pr}\left[ X_i = x \right]$. \\
+\end{dfn}
+
+\begin{lem}
+If $X_1, ... X_n$ are pairwise independent and $X = \sum X_{i=1}^n X_i$ then $\sigma^2_X = \sum_{i=1}^n \sigma_{X_i}^2$
+\end{lem}
+
+Given an algorithm $A(x,r)$ for deciding $L$ where $x \in L$ and $r$ is a random number in $\mathbb{Z}_n = \left\{ 0, \hdots, n-1 \right\}$ where $n$ is a prime s.t.
+\begin{align*}
+&\text{1) If } x \not\in L \text{ then } A(x,r) = 0 \; \forall r \in \mathbb{Z}_n \\
+&\text{2) If } x \in L \text{ then } A(x,r) = 1 \text{ for at least half of the choices of } r \in \mathbb{Z}_n
+\end{align*}
+Simple method: Pick $t$ numbers, $r_i, \hdots, r_t$ independently and uniformly at random from $\mathbb{Z}_n$.
+Output $\lor_{i=1}^t A(x,r_i)$.
+Given $x \in L$ the failure probability $\leq 1/2^t$.
+For $t=2$, the failure probability $\leq 1/4$.
+
+Two-point method: Now let $a,b \in \mathbb{Z}_n$ be picked uniformly at random.
+Let $r_i = (a_i + b) \mod n$, $Y_i = A(x, r_i)$ for $i = 0, \hdots, n-1$ and $Y=\sum_{i=0}^{n-1}Y_i$.
+We wish to bound $\mathbf{Pr}\left[ Y=0 \right]$ using Chebyshev's inequality.
+Now $\mu_{Y} = \sum^{n-1}_{i=0}\mu_{Y_i} \geq n/2$ and $\sigma_Y^2 = \sum_{i=0}^{n-1} \sigma^2_{Y_{i}} \geq n/4 \Rightarrow \sigma_Y \geq \sqrt{n}/2$.
+Inserting into Chebyshev's Inequality yields
+\begin{align*}
+\mathbf{Pr}\left[ Y = 0 \right] \leq \mathbf{Pr}\left[ \left| Y - \mu_Y \right| \geq \sqrt{n} \sigma_Y \right] \leq \frac{1}{n}
+\end{align*}
+which is better than $1/4$ for $n \geq 4$.
+
+\subsection{Coupon Collector's Problem}
+Given $n$ types of coupons, a trial means picking a coupon independently and at random.
+Let $X = \text{\# of trials to get all } n \text{ coupon types}$.
+For $i=0, \hdots, n-1$, the $i$th epoch consists of the trials starting just after the $i$th success and ending in the $(i+1)$th success.
+Define $X_i = \text{\# of trials in the } i\text{th epoch.}$, then $X=\sum_{i=0}^{n-1} X_i$.
+Now
+\begin{align*}
+\mathbf{E}\left[ X \right] = \sum_{i=0}^{n-1} \mathbf{E}\left[ X_i \right] = \sum_{i=0}^{n-1} \frac{n}{n-i} = \sum_{i=1}^{n} \frac{n}{i} =  n \sum_{i=1}^{n} \frac{1}{i} = n H_{n} = n \ln n + O(n)
+\end{align*}
+Let $\varepsilon_i^r$ denote the event that coupon type $i$ is not picked in the first $r$ trials. Then
+\begin{align*}
+\mathbf{Pr}\left[ \varepsilon_i^r \right] = \left(1-\frac{1}{n}\right)^r \leq e^{-r/n}
+\end{align*}
+Pick $r=\beta n \ln n$, where $\beta$ is a constant, then $\mathbf{Pr}\left[ \varepsilon_i^r \right] \leq e^{-\beta \ln n} = n^{-\beta}$ and
+\begin{align*}
+\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq \sum_{i=1}^n \mathbf{Pr}\left[ \varepsilon_i^r \right] \leq n^{1-\beta}
+\end{align*}
+For $\beta=2$, we get $\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq 1/n$.
+
+\subsection{Chernoff Bounds}
+Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$.
+Let $X=\sum_{i=1}^n X_i$ and $\mu = \mathbf{E} \left[ X \right] = \sum_{i=1}^n p_i$.
+Then
+\begin{align*}
+\mathbf{Pr}\left[ X > (1-\delta)\mu_X \right]
+&= \mathbf{Pr}\left[ e^{tX} > e^{t(1-\delta)\mu_X} \right]
+< \frac{\mathbf{E}\left[ e^{tX} \right]}{e^{t(1-\delta)\mu_X}}
+= \frac{\prod_{i=1}^n \mathbf{E}\left[ e^{tX_i} \right]}{e^{t(1-\delta)\mu_X}} \\
+&= \frac{\prod_{i=1}^n \left( p_i e^t + 1-p_i \right)}{e^{t(1-\delta)\mu_X}}
+= \frac{\prod_{i=1}^n \left( 1 + p_i (e^t - 1) \right)}{e^{t(1-\delta)\mu_X}} \\
+&\leq \frac{\prod_{i=1}^n e^{p_i (e^t-1)}}{e^{t(1-\delta)\mu_X}}
+\leq \left( \frac{e^{(e^t-1)}}{e^{t(1-\delta)}} \right)^{\mu_X}
+= \left( \frac{e^{\delta}}{\left( 1 + \delta \right)^{1+\delta}} \right)^{\mu_X}
+\end{align*}
+Furthermore
+\begin{align*}
+\mathbf{Pr}\left[ X < (1-\delta)\mu_X \right]
+&= \mathbf{Pr}\left[ -X < -(1-\delta)\mu_X \right]
+= \mathbf{Pr}\left[ e^{-tX} < e^{-t(1-\delta)\mu_X} \right] \\
+&< \left( \frac{e^{-\delta}}{\left( 1-\delta \right)^{1-\delta}} \right)^\mu
+= \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right) \ln (1-\delta)}} \right)^\mu \\
+&< \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right)\left( -\delta - \frac{\delta^2}{2} \right)}} \right)^\mu
+=  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2 + \frac{\delta^3}{2}}  } \right)^\mu \\
+&<  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2}  } \right)^\mu = \left( \frac{1}{e^{-\frac{\delta^2}{2} + \delta^2}  } \right)^\mu
+= \left( e^{\frac{\delta^2}{2} - \delta^2} \right)^\mu \\
+&= e^{\mu\frac{\delta^2}{2} - \mu\delta^2} = e^{-\mu\sigma^2/2}
+\end{align*}
+
+\subsection{Set-Balancing}
+Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t.
+$|| A\overline{b} ||_{\infty}$ is minimized.
+Pick each entry of $\overline{b}$ to be $1$ with probability $1/2$ and $-1$ otherwise.
+Now, $\mathbf{E} \left[ || A\overline{b} ||_{\infty} \right] = 0$.
+Consider a row $i$, $A_{i}$, of $A$, and assume for now that all entries are 1.
+Let $X$ be the \# of 1's in $\overline{b}$.
+Now
+\begin{align*}
+\mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] &\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \; \lor \; X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
+&\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \right] + \mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
+&\leq \frac{1}{n^2} + \frac{1}{n^2} = \frac{2}{n^2}
+\end{align*}
+and the third inequality follows from
+\begin{align*}
+\mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right]
+&= \mathbf{Pr} \left[ X < \left( 1 - \frac{2\sqrt{2n \ln n}}{n} \right)\frac{n}{2} \right]
+< e^{-\frac{n}{2} \left( \frac{2\sqrt{2n \ln n}}{n}\right)^2\frac{1}{2}} \\
+&< e^{-\frac{n}{4} \left( \frac{8n \ln n}{n^2}\right)} = e^{-2\ln n} = 1/n^2
+\end{align*}
+where the first inequality follows from applying the second Chernoff bound to the inequality with $\delta=\frac{2\sqrt{2n \ln n}}{n}$ and $\mu=n/2$.
+Finally applying an union bound over all rows we get
+\begin{align*}
+\mathbf{Pr} \left[ || A\overline{b} ||_{\infty} > 2\sqrt{2n \ln n} \right] \geq \sum_{i=1}^n \mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] \leq \sum_{i=1}^n \frac{2}{n^2} = \frac{2}{n}
+\end{align*}
 

### 2017-05-08 assignments/2/paper.tex

@@
 \section*{Exercise ???}
+
+\subsection*{3.1}
+We define $ee(m)$ to be the expected number of empty bins when m balls are thrown into n bins. We claim that $$ee(m)=\frac{(n-1)^m}{n^{m-1}}$$ and prove this by induction on the number m:\\
+\\
+First the base case. When 0 balls are thrown into n bins we of course expect to have n empty bins. We have $$ee(0)=\frac{(n-1)^0}{n^{0-1}}=\frac{1}{n^{-1}}=n$$ so the base case holds.\\
+\\
+Now for the induction step we assume that $ee(m)=\frac{(n-1)^m}{n^{m-1}}$ and need to prove $ee(m)=\frac{(n-1)^{m+1}}{n^m}$.
+Since the number of empty bins after one more ball can only be the same or one less than before that ball, we get that the expected number of empty bins after $m+1$ balls are equal to the expected number of empty bins after $m$ balls minus the probability that we get 1 less empty bin.
+So we get \begin{align*}ee(m+1)&=ee(m)-\frac{ee(m)}{n}=\frac{(n-1)^m}{n^{m-1}}-\frac{\left(\frac{(n-1)^m}{n^{m-1}}\right)}{n}\\&=\frac{n(n-1)^m}{n^m}-\frac{(n-1)^m}{n^m}=\frac{(n-1)(n-1)^m}{n^m}=\frac{(n-1)^{m+1}}{n^m}\end{align*}
+And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.\\
+\\
+For $m=n$ we get $$ee(n)=\frac{(n-1)^n}{n^{n-1}}=n\frac{(n-1)^n}{n^n}=n\left(\frac{n-1}{n}\right)^n=n\left(n-\frac{1}{n}\right)^n$$ And since $\left(n-\frac{1}{n}\right)^n$ goes towards $e$ when n goes towards infinity, we get that for large n the expected number of empty bins approaches $\frac{n}{e}$ as we wanted to show.
 

### 2017-05-08 assignments/2/paper.tex
Update paper.tex

--- 2017-05-07 assignments/2/paper.tex
@@
 
-\section*{Exercise ???}
+\section*{Problems}
+
+\subsection*{2.1}
+We prove by induction over k that for any deterministic evaluation algorithm, there are instances of $T_{d,k}$ that forces the algorithm to read the values on all $d^{2k}$ leaves and the last leaf read determines whether the root is 1 or 0.\\
+\\
+Base step: for k=0 we only have one leaf, so any deterministic evaluation algorithm has to read them all. Furthermore that leaf is the root, so it determines whether the root is 1 or 0.\\
+\\
+Induction step: We assume it holds for any tree with size $k$, and need to show that it holds for any tree of size $k+1$.
+For a tree of size $k+1$, we have at the top a root that is an AND-note.
+Below that we have two OR-nodes, and below that we have 4 trees of size k.
+Since a deterministic evaluation algorithm looks at the nodes in a deterministic order, it also looks at the nodes in each of the subtrees in a specific order, and we can use the induction hypothesis.\\
+There has to be one of the subtrees that are the first to get its last node inspected.
+We use the induction hypothesis and set the last node inspected in that subtree in the way that makes the root of that subtree 0.
+Since the parent to that root is an OR-node, we don't know yet whether it will be 1 or 0.\\
+The 2nd subtree to get its last node inspected has 2 possibilities: Its root either shares a parent with the subtree we have already evaluated or it doesn't.\\
+In the case where it shares a parent, we set the last node inspected in the way that makes the root 1.
+The parent OR-node can now look at a 1 and a 0, so the result is 1, but then we don't know the result of the AND-node in the root of the whole tree.
+We set the last node in the 3rd subtree to get fully evaluated in such a way that the root of that subtree is 0, and we have to inspect the 4th subtree.
+We can set the last node in the 4th subtree both so the root of the whole tree is 0 and so the root of the whole tree is 1.
+In the case, where the 2nd subtree fully evaluated does not share a parent with the first, we set the last node in that subtree, so the root of the tree is 0, and we can't evaluate the parent OR-node.
+The 3rd subtree fully evaluated now either shares a parent with the first or the second subtree fully evaluated.
+We set the last node in the 3rd subtree in such a way that the root is 1, and we then have to also fully evaluate the 4th subtree before we can evaluate the root of the whole tree.
+Depending what we set the last node inspected in the 4th subtree, the root of the whole tree can both result in 0 or 1.\\
+And we're done.
 
 \subsection*{3.1}
-We define $ee(m)$ to be the expected number of empty bins when m balls are thrown into n bins. We claim that $$ee(m)=\frac{(n-1)^m}{n^{m-1}}$$ and prove this by induction on the number m:\\
-\\
-First the base case. When 0 balls are thrown into n bins we of course expect to have n empty bins. We have $$ee(0)=\frac{(n-1)^0}{n^{0-1}}=\frac{1}{n^{-1}}=n$$ so the base case holds.\\
-\\
-Now for the induction step we assume that $ee(m)=\frac{(n-1)^m}{n^{m-1}}$ and need to prove $ee(m)=\frac{(n-1)^{m+1}}{n^m}$.
-Since the number of empty bins after one more ball can only be the same or one less than before that ball, we get that the expected number of empty bins after $m+1$ balls are equal to the expected number of empty bins after $m$ balls minus the probability that we get 1 less empty bin.
-So we get \begin{align*}ee(m+1)&=ee(m)-\frac{ee(m)}{n}=\frac{(n-1)^m}{n^{m-1}}-\frac{\left(\frac{(n-1)^m}{n^{m-1}}\right)}{n}\\&=\frac{n(n-1)^m}{n^m}-\frac{(n-1)^m}{n^m}=\frac{(n-1)(n-1)^m}{n^m}=\frac{(n-1)^{m+1}}{n^m}\end{align*}
-And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.\\
-\\
-For $m=n$ we get $$ee(n)=\frac{(n-1)^n}{n^{n-1}}=n\frac{(n-1)^n}{n^n}=n\left(\frac{n-1}{n}\right)^n=n\left(n-\frac{1}{n}\right)^n$$ And since $\left(n-\frac{1}{n}\right)^n$ goes towards $e$ when n goes towards infinity, we get that for large n the expected number of empty bins approaches $\frac{n}{e}$ as we wanted to show.
+Let $X$ be the number of empty bins after $m$ balls and $X_i$ be 1 if bin $i$ is empty after $m$ balls and 0 otherwise for $0\leq i\leq n$.
+For $X_i$ to be empty after m balls, we need all the m balls to land in a different bin, so we have $Pr[X_i]=\left(1-\frac{1}{n}\right)^m$ and $$E[X]=\Sigma_{i=0}^nE[X_i]=\Sigma_{i=0}^nPr[X_i]=\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^m=n\left(1-\frac{1}{n}\right)^m$$
+For $m=n$ we get $$E[X]=n\left(1-\frac{1}{n}\right)^n\approx\frac{n}{e}$$
+
+\subsection*{3.3(a)}
+Let $X$ be the number of memory modules that has received exactly 1 memory request after $n$ memory requests has been sent and $X_i$ be 1 if memory module $i$ has received exactly 1 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$.
+We have $Pr[X_i]={n\choose 1}\left(\frac{1}{n}\right)^1\left(1-\frac{1}{n}\right)^{n-1}=\left(1-\frac{1}{n}\right)^{n-1}$.\\
+Let $Y$ be the number of memory modules that has received exactly 2 memory request after $n$ memory requests has been sent and $Y_i$ be 1 if memory module $i$ has received exactly 2 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$.
+We have $Pr[Y_i]={n\choose 2}\left(\frac{1}{n}\right)^2\left(1-\frac{1}{n}\right)^{n-2}={n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}$.
+Let $Z$ be the number of processors whose requests are satisfied.
+We have \begin{align*}E[Z]&=E[X]+2E[Y]=\left(\Sigma_{i=0}^nE[X_i]\right)+2\Sigma_{i=0}^nE[Y_i]=\left(\Sigma_{i=0}^nPr[X_i]\right)+2\Sigma_{i=0}^nPr[Y_i]\\&=\left(\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^{n-1}\right)+2\Sigma_{i=0}^n{n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}\\&=n\left(1-\frac{1}{n}\right)^{n-1}+2n\frac{n(n-1)}{2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}=\frac{n}{1-\frac{1}{n}}\left(1-\frac{1}{n}\right)^n+(n-1)\left(1-\frac{1}{n}\right)^{n-2}\\&=\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n+\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n\approx\frac{2e(n-1)}{\left(\frac{n-1}{n}\right)^2}=\frac{2en^2}{n-1}\end{align*}
+This is around $2en$, which is of course impossible (we can satisfy more processor than are available), so we must have made a mistake somewhere.
 

### 2017-05-11 assignments/2/paper.tex
Assignment 2 summary.

--- 2017-05-08 assignments/2/paper.tex
@@
 
-\section*{Exercise ???}
+\section*{Assigment}
 
@@
 
-\section{Summary}
-\subsection{Balls into bins}
+\section*{Summary}
+\subsection*{Balls into bins}
 Given $n$ balls and $n$ bins, and ball $i$ goes into bin $j$ with probability $1/n$.
@@
 
-\subsection{Markov's Inequality}
+\subsection*{Markov's Inequality}
 Given a non-negative random variable $Y$ and $t > 0$, then
@@
 
-\subsection{Chebyshev's Inequality}
+\subsection*{Chebyshev's Inequality}
 Given rand.
@@
 
-\subsection{Lazy Selection}
+\subsection*{Lazy Selection}
 Refer to the algorithm given in the book.
@@
 
-\subsection{Two-point Sampling}
+\subsection*{Two-point Sampling}
 \begin{lem}
@@
 
-\subsection{Coupon Collector's Problem}
+\subsection*{Coupon Collector's Problem}
 Given $n$ types of coupons, a trial means picking a coupon independently and at random.
@@
 
-\subsection{Chernoff Bounds}
+\subsection*{Chernoff Bounds}
 Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$.
@@
 
-\subsection{Set-Balancing}
+\subsection**{Set-Balancing}
 Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t.

### 2017-05-15 assignments/2/paper.tex

@@
 
-\section*{Assigment}
+\section*{Problems}
+
+\subsection*{2.1}
+We prove by induction over k that for any deterministic evaluation algorithm, there are instances of $T_{d,k}$ that forces the algorithm to read the values on all $d^{2k}$ leaves and the last leaf read determines whether the root is 1 or 0.\\
+\\
+Base step: for k=0 we only have one leaf, so any deterministic evaluation algorithm has to read them all. Furthermore that leaf is the root, so it determines whether the root is 1 or 0.\\
+\\
+Induction step: We assume it holds for any tree with size $k$, and need to show that it holds for any tree of size $k+1$.
+For a tree of size $k+1$, we have at the top a root that is an AND-note.
+Below that we have two OR-nodes, and below that we have 4 trees of size k.
+Since a deterministic evaluation algorithm looks at the nodes in a deterministic order, it also looks at the nodes in each of the subtrees in a specific order, and we can use the induction hypothesis.\\
+There has to be one of the subtrees that are the first to get its last node inspected.
+We use the induction hypothesis and set the last node inspected in that subtree in the way that makes the root of that subtree 0.
+Since the parent to that root is an OR-node, we don't know yet whether it will be 1 or 0.\\
+The 2nd subtree to get its last node inspected has 2 possibilities: Its root either shares a parent with the subtree we have already evaluated or it doesn't.\\
+In the case where it shares a parent, we set the last node inspected in the way that makes the root 1.
+The parent OR-node can now look at a 1 and a 0, so the result is 1, but then we don't know the result of the AND-node in the root of the whole tree.
+We set the last node in the 3rd subtree to get fully evaluated in such a way that the root of that subtree is 0, and we have to inspect the 4th subtree.
+We can set the last node in the 4th subtree both so the root of the whole tree is 0 and so the root of the whole tree is 1.
+In the case, where the 2nd subtree fully evaluated does not share a parent with the first, we set the last node in that subtree, so the root of the tree is 0, and we can't evaluate the parent OR-node.
+The 3rd subtree fully evaluated now either shares a parent with the first or the second subtree fully evaluated.
+We set the last node in the 3rd subtree in such a way that the root is 1, and we then have to also fully evaluate the 4th subtree before we can evaluate the root of the whole tree.
+Depending what we set the last node inspected in the 4th subtree, the root of the whole tree can both result in 0 or 1.\\
+And we're done.
 
 \subsection*{3.1}
-We define $ee(m)$ to be the expected number of empty bins when m balls are thrown into n bins. We claim that $$ee(m)=\frac{(n-1)^m}{n^{m-1}}$$ and prove this by induction on the number m:\\
-\\
-First the base case. When 0 balls are thrown into n bins we of course expect to have n empty bins. We have $$ee(0)=\frac{(n-1)^0}{n^{0-1}}=\frac{1}{n^{-1}}=n$$ so the base case holds.\\
-\\
-Now for the induction step we assume that $ee(m)=\frac{(n-1)^m}{n^{m-1}}$ and need to prove $ee(m)=\frac{(n-1)^{m+1}}{n^m}$.
-Since the number of empty bins after one more ball can only be the same or one less than before that ball, we get that the expected number of empty bins after $m+1$ balls are equal to the expected number of empty bins after $m$ balls minus the probability that we get 1 less empty bin.
-So we get \begin{align*}ee(m+1)&=ee(m)-\frac{ee(m)}{n}=\frac{(n-1)^m}{n^{m-1}}-\frac{\left(\frac{(n-1)^m}{n^{m-1}}\right)}{n}\\&=\frac{n(n-1)^m}{n^m}-\frac{(n-1)^m}{n^m}=\frac{(n-1)(n-1)^m}{n^m}=\frac{(n-1)^{m+1}}{n^m}\end{align*}
-And we have proven the induction step, so by induction we get for any m $ee(m)=\frac{(n-1)^m}{n^{m-1}}$.\\
-\\
-For $m=n$ we get $$ee(n)=\frac{(n-1)^n}{n^{n-1}}=n\frac{(n-1)^n}{n^n}=n\left(\frac{n-1}{n}\right)^n=n\left(n-\frac{1}{n}\right)^n$$ And since $\left(n-\frac{1}{n}\right)^n$ goes towards $e$ when n goes towards infinity, we get that for large n the expected number of empty bins approaches $\frac{n}{e}$ as we wanted to show.
+Let $X$ be the number of empty bins after $m$ balls and $X_i$ be 1 if bin $i$ is empty after $m$ balls and 0 otherwise for $0\leq i\leq n$.
+For $X_i$ to be empty after m balls, we need all the m balls to land in a different bin, so we have $Pr[X_i]=\left(1-\frac{1}{n}\right)^m$ and $$E[X]=\Sigma_{i=0}^nE[X_i]=\Sigma_{i=0}^nPr[X_i]=\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^m=n\left(1-\frac{1}{n}\right)^m$$
+For $m=n$ we get $$E[X]=n\left(1-\frac{1}{n}\right)^n\approx\frac{n}{e}$$
+
+\subsection*{3.3(a)}
+Let $X$ be the number of memory modules that has received exactly 1 memory request after $n$ memory requests has been sent and $X_i$ be 1 if memory module $i$ has received exactly 1 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$.
+We have $Pr[X_i]={n\choose 1}\left(\frac{1}{n}\right)^1\left(1-\frac{1}{n}\right)^{n-1}=\left(1-\frac{1}{n}\right)^{n-1}$.\\
+Let $Y$ be the number of memory modules that has received exactly 2 memory request after $n$ memory requests has been sent and $Y_i$ be 1 if memory module $i$ has received exactly 2 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$.
+We have $Pr[Y_i]={n\choose 2}\left(\frac{1}{n}\right)^2\left(1-\frac{1}{n}\right)^{n-2}={n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}$.
+Let $Z$ be the number of processors whose requests are satisfied.
+We have \begin{align*}E[Z]&=E[X]+2E[Y]=\left(\Sigma_{i=0}^nE[X_i]\right)+2\Sigma_{i=0}^nE[Y_i]=\left(\Sigma_{i=0}^nPr[X_i]\right)+2\Sigma_{i=0}^nPr[Y_i]\\&=\left(\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^{n-1}\right)+2\Sigma_{i=0}^n{n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}\\&=n\left(1-\frac{1}{n}\right)^{n-1}+2n\frac{n(n-1)}{2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}=\frac{n}{1-\frac{1}{n}}\left(1-\frac{1}{n}\right)^n+(n-1)\left(1-\frac{1}{n}\right)^{n-2}\\&=\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n+\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n\approx\frac{2e(n-1)}{\left(\frac{n-1}{n}\right)^2}=\frac{2en^2}{n-1}\end{align*}
+This is around $2en$, which is of course impossible (we can satisfy more processor than are available), so we must have made a mistake somewhere.
 
@@
 
-\subsection**{Set-Balancing}
+\subsection*{Set-Balancing}
 Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t.

### 2017-05-15 assignments/2/paper.tex
Assignment 2 re-handin. Assignment 3 stub + 1/2 summary.

@@
 
-\section*{Summary}
+\section*{Summary - April 27th and May 2nd}
+\subsection*{Game tree evaluation}
+Consider an algorithm for evaluating perfect, boolean, binary game trees with $n$ $0/1$-leaves consisting of AND-nodes (min) on even levels and OR-nodes (max) on uneven levels. Then
+\\
+\begin{thm}
+Let $k$ denote the number of and-or levels in the tree, so that the height of the tree is $2k$ and the number of leaves is $n=4^k$.
+Then any det.
+algorithm might need to inspect all leaves and thus runs in $\Omega(n)$ or $\Omega(4^k)$ time in the worst case.
+\end{thm}
+We will give a rand.
+algorithm running in expected time $O(3^k) \approx O(n^{0.793})$.
+Let $T_w$ denote the subtree rooted at $w$ and $N_w$ be the number of leaves in $T_w$ inspected by the random.
+algorithm.
+Then we wish to show $\mathbf{E}\left[ N_r \right] \leq 3^k$ where $r$ is the root of such a game tree and do so by induction on $k \geq 0$.
+When $k=0$, $\mathbf{E}\left[ N_r \right] = 3^0 = 1 \leq 3^k$.
+Now for $k > 0$ we have
+\begin{figure}[!ht]
+\includegraphics[width=8cm]{img/gametree.png}
+\centering
+\end{figure}
+
+Consider child $u$ of $w$.
+Then in case $1$, both children are $0$ and the parent is an OR-node.
+Thus $\mathbf{E}\left[ N_u \right] \leq 2 \cdot 3^{k-1}$.
+In case $2$, one of the children are $1$ and another is $0$ and $\mathbf{E}\left[ N_u \right] \leq 3^{k-1} + \frac{1}{2} \cdot 3^{k-1} = \frac{3}{2} \cdot 3^{k-1}$.
+Next, consider the root node $w$.
+In case $1$, one child node is $0$ and the other is $1$ and $\mathbf{E}\left[ N_w \right] \leq \frac{1}{2} \cdot \left( 2 \cdot 3^{k-1} \right) + \frac{1}{2} \cdot \left( \frac{3}{2} \cdot 3^{k-1} + 2 \cdot 3^{k-1} \right) < 3^k$.
+In case $2$, both child nodes return $1$ and $\mathbf{E}\left[ N_w \right] \leq 2 \cdot \frac{3}{2} 3^{k-1} = 3^{k}$.
+
+\subsection*{Game theory (von Neumann)}
+Let $M$ be an $n \times n$ payoff-matrix in which each entry $M_{ij}$ denotes the payoff to the row player from the column player when the row player chooses strategy $i$ and the column player chooses strategy $j$.
+\\
+\begin{dfn}
+(Pure strategies) Define the values for row and column-player as $V_R = \max\limits_{i} \min\limits_{j} M_{ij}$ and $V_C = \min\limits_{j} \max\limits_{i} M_{ij}$, respectively
+\end{dfn}
+
+In general $V_R \leq V_C$ and when $V_R=V_C$ the game has a solution.
+\\
+\begin{dfn}
+(Mixed strategies) Strategy for row player is a prob.
+dist.
+over the rows of $M_{ij}$, $\overline{p}=\left( p_1, \hdots, p_n \right)$ and similarly for the column player with $\overline{q}$ over the columns in place of $\overline{p}$.
+Define the values for the row-player as $V_R = \max\limits_{\overline{p}} \min\limits_{\overline{q}} \sum \sum p_i \cdot q_j \cdot M_{ij}$ and $V_C = \min\limits_{\overline{q}} \max\limits_{\overline{p}} \sum \sum p_i \cdot q_j \cdot M_{ij}$
+\end{dfn}
+The following theorem shows that mixed-strategy games always have a solution \\
+\begin{thm}
+(von Neumann) For any game and any mixed strategies $\overline{p}$ and $\overline{q}$, $V_R = V_C$
+\end{thm}
+And the following theorem shows that the optimal mixed strategy, given the other player chooses a pure strategy always results in an equal payoff \\
+\begin{thm}
+(Loomis) $\max\limits_{\overline{p}} \min\limits_j \sum p_i M_{ij} = \min\limits_{\overline{q}} \max\limits_i \sum q_j M_{ij} $
+\end{thm}
+
+\subsection*{Yao's Minimax Principle}
+Given a finite set $\mathcal{I}$ of problem instances all of the same size and a finite set, $\mathcal{A}$, of det.
+algorithms for $\mathcal{I}$, let $C(I,A)$ denote the cost associated with running $A \in \mathcal{A}$ on a problem instance $I \in \mathcal{I}$
+\begin{figure}[!ht]
+\includegraphics[width=8cm]{img/payoffmatrix}
+\centering
+\end{figure}
+\begin{thm}
+(Yao's Minimax Principle) Given dists.
+$\overline{p}$ and $\overline{q}$, let $I_{\overline{p}}$ be picked from $\mathcal{I}$ according to $\overline{p}$, dually for $A_{\overline{q}}$.
+Then $\min\limits_{A \in \mathcal{A}} \mathbf{E}\left[ C(I_{\overline{p}}, A) \right] \leq \max\limits_{I \in \mathcal{I}} \mathbf{E} \left[ C(I, A_{\overline{q}}) \right]$
+\paragraph{Proof:} $\min\limits_{A \in \mathcal{A}} \mathbf{E}\left[ C(I_{\overline{p}}, A) \right] = \min\limits_{A \in \mathcal{A}}\limits \sum\limits_{I \in \mathcal{I}} \overline{p}(I) \cdot C(I_{\overline{p}}, A) \leq \sum\limits_{A \in \mathcal{A}} \overline{q}(A) \sum\limits_{I \in \mathcal{I}} \overline{p}(I) \cdot C(I_{\overline{p}}, A) = \sum\limits_{I \in \mathcal{I}} \overline{p}(I) \sum\limits_{A \in \mathcal{A}} \overline{q}(A) \cdot C(I_{\overline{p}}, A) \leq \max\limits_{I \in \mathcal{I}} \mathbf{E} \left[ C(I, A_{\overline{q}}) \right]$
+\end{thm}
+
+\subsection*{Lower bound for game tree evaluation}
+Choose $\overline{p}$: Independently, set each leaf to $1$ with prob. $p=\frac{3-\sqrt{5}}{2}$. $p$ satisfies $p=(1-p)^2$. Change tree to NOR tree. This does not change the semantics of the tree. \\
+\begin{lem}
+After fixing leaves according to $\overline{p}$, each node of $T$ is $1$ with prob. $p$.
+\paragraph{Proof:} $P[w\text{ set to } 1] = (1-p)^2 = p$.
+\end{lem}
+
+\subsection*{Derandomization}
+Consider boolean func. $f : \left\{ 0,1 \right\}^* \rightarrow \left\{ 0,1 \right\}$. A rand. boolean circuit, $C$, computes $f$ if:
+\begin{enumerate}
+\item $\forall x_1, \hdots, x_n$ s.t. $f(x_1, \hdots, x_n) = 0$, $C$ outputs $0$.
+\item $\forall x_1, \hdots, x_n$ s.t. $f(x_1, \hdots, x_n) = 1$, $C$ outputs $1$ for at least half of the $2^m$ assignments of values to $r_1, \hdots, r_m$.\\
+\end{enumerate} 
+\begin{dfn}
+(Rand.) circuits $C_1, \hdots$ is a (rand.) circuit family for $f$ if $C_i$ computes $f$ for all $x_1, \hdots, x_i$. \\
+\end{dfn}
+
+\begin{dfn}
+A circuit family $C_1, \hdots$ has polynomial size if $\forall n \left| C_n \right| \leq n^c$ where $c$ is a constant. \\
+\end{dfn}
+
+\begin{thm}
+If $f$ has a polynomial sized random circuit family then $f$ has a polynomial sized deterministic circuit family.
+
+\paragraph{Proof:} Given $n$, for each $C_n$, construct table $T$ with a row for each $2^n$ choices of $x_1, \hdots, x_n$ and a column for each $2^m$ choices of $r_1, \hdots, r_m$.
+Now if $x \not\in L$ then $\forall r, C_n(x,r) = 0$ and if $x \in L$ then $C_{n}(x,r) = 1$ for at least half the choices of $r$.
+\begin{enumerate}
+\item Remove all rows where $f$ only outputs 0.
+\item For each remaining row, at least half of the entries are 1, meaning that there must exist a column where at least half of the entries are 1.
+Construct circuit $T_1$ by hardwiring $r_1, \hdots, r_m$ to values for that column.
+\item Remove rows having a 1 in the chosen column.
+\item Construct $T_2$, $T_3, \hdots, T_k$ by repeating steps 2 and 3.
+\item Output the OR of the circuits $T_1, \hdots, T_k$
+\end{enumerate}
+\end{thm}
+We wish to bound the number of circuits and note that in each iteration we remove at least half of the rows of $T$.
+Hence $k \leq \log(2^n) = n$ the number of circuits in the deterministic circuit family is polynomial.
+
 \subsection*{Balls into bins}
@@
 &= \mathbf{Pr}\left[ Y \geq t^2\mathbf{E}\left[Y\right] \right] \\
-&= \frac{1}{t^2}
+&\leq \frac{1}{t^2}
 \end{align*}
@@
 
-\subsection*{Coupon Collector's Problem}
-Given $n$ types of coupons, a trial means picking a coupon independently and at random.
-Let $X = \text{\# of trials to get all } n \text{ coupon types}$.
-For $i=0, \hdots, n-1$, the $i$th epoch consists of the trials starting just after the $i$th success and ending in the $(i+1)$th success.
-Define $X_i = \text{\# of trials in the } i\text{th epoch.}$, then $X=\sum_{i=0}^{n-1} X_i$.
-Now
-\begin{align*}
-\mathbf{E}\left[ X \right] = \sum_{i=0}^{n-1} \mathbf{E}\left[ X_i \right] = \sum_{i=0}^{n-1} \frac{n}{n-i} = \sum_{i=1}^{n} \frac{n}{i} =  n \sum_{i=1}^{n} \frac{1}{i} = n H_{n} = n \ln n + O(n)
-\end{align*}
-Let $\varepsilon_i^r$ denote the event that coupon type $i$ is not picked in the first $r$ trials. Then
-\begin{align*}
-\mathbf{Pr}\left[ \varepsilon_i^r \right] = \left(1-\frac{1}{n}\right)^r \leq e^{-r/n}
-\end{align*}
-Pick $r=\beta n \ln n$, where $\beta$ is a constant, then $\mathbf{Pr}\left[ \varepsilon_i^r \right] \leq e^{-\beta \ln n} = n^{-\beta}$ and
-\begin{align*}
-\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq \sum_{i=1}^n \mathbf{Pr}\left[ \varepsilon_i^r \right] \leq n^{1-\beta}
-\end{align*}
-For $\beta=2$, we get $\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq 1/n$.
-
-\subsection*{Chernoff Bounds}
-Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$.
-Let $X=\sum_{i=1}^n X_i$ and $\mu = \mathbf{E} \left[ X \right] = \sum_{i=1}^n p_i$.
-Then
-\begin{align*}
-\mathbf{Pr}\left[ X > (1-\delta)\mu_X \right]
-&= \mathbf{Pr}\left[ e^{tX} > e^{t(1-\delta)\mu_X} \right]
-< \frac{\mathbf{E}\left[ e^{tX} \right]}{e^{t(1-\delta)\mu_X}}
-= \frac{\prod_{i=1}^n \mathbf{E}\left[ e^{tX_i} \right]}{e^{t(1-\delta)\mu_X}} \\
-&= \frac{\prod_{i=1}^n \left( p_i e^t + 1-p_i \right)}{e^{t(1-\delta)\mu_X}}
-= \frac{\prod_{i=1}^n \left( 1 + p_i (e^t - 1) \right)}{e^{t(1-\delta)\mu_X}} \\
-&\leq \frac{\prod_{i=1}^n e^{p_i (e^t-1)}}{e^{t(1-\delta)\mu_X}}
-\leq \left( \frac{e^{(e^t-1)}}{e^{t(1-\delta)}} \right)^{\mu_X}
-= \left( \frac{e^{\delta}}{\left( 1 + \delta \right)^{1+\delta}} \right)^{\mu_X}
-\end{align*}
-Furthermore
-\begin{align*}
-\mathbf{Pr}\left[ X < (1-\delta)\mu_X \right]
-&= \mathbf{Pr}\left[ -X < -(1-\delta)\mu_X \right]
-= \mathbf{Pr}\left[ e^{-tX} < e^{-t(1-\delta)\mu_X} \right] \\
-&< \left( \frac{e^{-\delta}}{\left( 1-\delta \right)^{1-\delta}} \right)^\mu
-= \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right) \ln (1-\delta)}} \right)^\mu \\
-&< \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right)\left( -\delta - \frac{\delta^2}{2} \right)}} \right)^\mu
-=  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2 + \frac{\delta^3}{2}}  } \right)^\mu \\
-&<  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2}  } \right)^\mu = \left( \frac{1}{e^{-\frac{\delta^2}{2} + \delta^2}  } \right)^\mu
-= \left( e^{\frac{\delta^2}{2} - \delta^2} \right)^\mu \\
-&= e^{\mu\frac{\delta^2}{2} - \mu\delta^2} = e^{-\mu\sigma^2/2}
-\end{align*}
-
-\subsection*{Set-Balancing}
-Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t.
-$|| A\overline{b} ||_{\infty}$ is minimized.
-Pick each entry of $\overline{b}$ to be $1$ with probability $1/2$ and $-1$ otherwise.
-Now, $\mathbf{E} \left[ || A\overline{b} ||_{\infty} \right] = 0$.
-Consider a row $i$, $A_{i}$, of $A$, and assume for now that all entries are 1.
-Let $X$ be the \# of 1's in $\overline{b}$.
-Now
-\begin{align*}
-\mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] &\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \; \lor \; X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
-&\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \right] + \mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
-&\leq \frac{1}{n^2} + \frac{1}{n^2} = \frac{2}{n^2}
-\end{align*}
-and the third inequality follows from
-\begin{align*}
-\mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right]
-&= \mathbf{Pr} \left[ X < \left( 1 - \frac{2\sqrt{2n \ln n}}{n} \right)\frac{n}{2} \right]
-< e^{-\frac{n}{2} \left( \frac{2\sqrt{2n \ln n}}{n}\right)^2\frac{1}{2}} \\
-&< e^{-\frac{n}{4} \left( \frac{8n \ln n}{n^2}\right)} = e^{-2\ln n} = 1/n^2
-\end{align*}
-where the first inequality follows from applying the second Chernoff bound to the inequality with $\delta=\frac{2\sqrt{2n \ln n}}{n}$ and $\mu=n/2$.
-Finally applying an union bound over all rows we get
-\begin{align*}
-\mathbf{Pr} \left[ || A\overline{b} ||_{\infty} > 2\sqrt{2n \ln n} \right] \geq \sum_{i=1}^n \mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] \leq \sum_{i=1}^n \frac{2}{n^2} = \frac{2}{n}
-\end{align*}
-
 \end{document}

### 2017-05-15 assignments/3/paper.tex
Assignment 2 re-handin. Assignment 3 stub + 1/2 summary.

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[]{algorithm2e}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Randomized Algorithms \\
       Assignment 3}

\author{Viktor Hansen \& Simon Rueskov Schleicher}

\maketitle

\begin{abstract}
  This is the third weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{Exercise 3.10}
We wish to find an upper bound on $\mathbf{Pr} \left[ X > \beta n \ln n \right] \leq \mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \beta n \ln n - \mu_X \right]$. Now, setting $t\sigma_X = \beta n \ln n - \mu_X$, we get $t = \frac{\beta n \ln n - \mu_X}{\sigma_X}$. Inserting and applying Chebyshev's inequality yields
\begin{align*}
\mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \frac{\beta n \ln n - \mu_X}{\sigma_X} \sigma_X \right] \leq \frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2}
\end{align*}
Inserting $\mu_X = \mathbf{E} \left[ X \right] = n \ln n + O(n)$ gives
\begin{align*}
\frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln n - n \ln n - cn)^2} = \frac{\sigma_X^2}{(n \ln n(\beta - 1) - cn)^2}
\end{align*}

\pagebreak

\section*{Summary}
\subsection*{Coupon Collector's Problem}
Given $n$ types of coupons, a trial means picking a coupon independently and at random. Let $X = \text{\# of trials to get all } n \text{ coupon types}$. For $i=0, \hdots, n-1$, the $i$th epoch consists of the trials starting just after the $i$th success and ending in the $(i+1)$th success. Define $X_i = \text{\# of trials in the } i\text{th epoch.}$, then $X=\sum_{i=0}^{n-1} X_i$. Now
\begin{align*}
\mathbf{E}\left[ X \right] = \sum_{i=0}^{n-1} \mathbf{E}\left[ X_i \right] = \sum_{i=0}^{n-1} \frac{n}{n-i} = \sum_{i=1}^{n} \frac{n}{i} =  n \sum_{i=1}^{n} \frac{1}{i} = n H_{n} = n \ln n + O(n)
\end{align*}
Let $\varepsilon_i^r$ denote the event that coupon type $i$ is not picked in the first $r$ trials. Then
\begin{align*}
\mathbf{Pr}\left[ \varepsilon_i^r \right] = \left(1-\frac{1}{n}\right)^r \leq e^{-r/n}
\end{align*}
Pick $r=\beta n \ln n$, where $\beta$ is a constant, then $\mathbf{Pr}\left[ \varepsilon_i^r \right] \leq e^{-\beta \ln n} = n^{-\beta}$ and
\begin{align*}
\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq \sum_{i=1}^n \mathbf{Pr}\left[ \varepsilon_i^r \right] \leq n^{1-\beta}
\end{align*}
For $\beta=2$, we get $\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq 1/n$.

\subsection*{Chernoff Bounds}
Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$. Let $X=\sum_{i=1}^n X_i$ and $\mu = \mathbf{E} \left[ X \right] = \sum_{i=1}^n p_i$. Then
\begin{align*}
\mathbf{Pr}\left[ X > (1-\delta)\mu_X \right]
&= \mathbf{Pr}\left[ e^{tX} > e^{t(1-\delta)\mu_X} \right]
< \frac{\mathbf{E}\left[ e^{tX} \right]}{e^{t(1-\delta)\mu_X}}
= \frac{\prod_{i=1}^n \mathbf{E}\left[ e^{tX_i} \right]}{e^{t(1-\delta)\mu_X}} \\
&= \frac{\prod_{i=1}^n \left( p_i e^t + 1-p_i \right)}{e^{t(1-\delta)\mu_X}}
= \frac{\prod_{i=1}^n \left( 1 + p_i (e^t - 1) \right)}{e^{t(1-\delta)\mu_X}} \\
&\leq \frac{\prod_{i=1}^n e^{p_i (e^t-1)}}{e^{t(1-\delta)\mu_X}}
\leq \left( \frac{e^{(e^t-1)}}{e^{t(1-\delta)}} \right)^{\mu_X}
= \left( \frac{e^{\delta}}{\left( 1 + \delta \right)^{1+\delta}} \right)^{\mu_X}
\end{align*}
Furthermore
\begin{align*}
\mathbf{Pr}\left[ X < (1-\delta)\mu_X \right]
&= \mathbf{Pr}\left[ -X < -(1-\delta)\mu_X \right]
= \mathbf{Pr}\left[ e^{-tX} < e^{-t(1-\delta)\mu_X} \right] \\
&< \left( \frac{e^{-\delta}}{\left( 1-\delta \right)^{1-\delta}} \right)^\mu
= \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right) \ln (1-\delta)}} \right)^\mu \\
&< \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right)\left( -\delta - \frac{\delta^2}{2} \right)}} \right)^\mu
=  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2 + \frac{\delta^3}{2}}  } \right)^\mu \\
&<  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2}  } \right)^\mu = \left( \frac{1}{e^{-\frac{\delta^2}{2} + \delta^2}  } \right)^\mu
= \left( e^{\frac{\delta^2}{2} - \delta^2} \right)^\mu \\
&= e^{\mu\frac{\delta^2}{2} - \mu\delta^2} = e^{-\mu\sigma^2/2}
\end{align*}

\subsection*{Set-Balancing}
Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t. $|| A\overline{b} ||_{\infty}$ is minimized. Pick each entry of $\overline{b}$ to be $1$ with probability $1/2$ and $-1$ otherwise. Now, $\mathbf{E} \left[ || A\overline{b} ||_{\infty} \right] = 0$. Consider a row $i$, $A_{i}$, of $A$, and assume for now that all entries are 1. Let $X$ be the \# of 1's in $\overline{b}$. Now
\begin{align*}
\mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] &\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \; \lor \; X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
&\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \right] + \mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
&\leq \frac{1}{n^2} + \frac{1}{n^2} = \frac{2}{n^2}
\end{align*}
and the third inequality follows from
\begin{align*}
\mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right]
&= \mathbf{Pr} \left[ X < \left( 1 - \frac{2\sqrt{2n \ln n}}{n} \right)\frac{n}{2} \right]
< e^{-\frac{n}{2} \left( \frac{2\sqrt{2n \ln n}}{n}\right)^2\frac{1}{2}} \\
&< e^{-\frac{n}{4} \left( \frac{8n \ln n}{n^2}\right)} = e^{-2\ln n} = 1/n^2
\end{align*}
where the first inequality follows from applying the second Chernoff bound to the inequality with $\delta=\frac{2\sqrt{2n \ln n}}{n}$ and $\mu=n/2$. Finally applying an union bound over all rows we get
\begin{align*}
\mathbf{Pr} \left[ || A\overline{b} ||_{\infty} > 2\sqrt{2n \ln n} \right] \geq \sum_{i=1}^n \mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] \leq \sum_{i=1}^n \frac{2}{n^2} = \frac{2}{n}
\end{align*}

\end{document}

### 2017-05-22 assignments/3/paper.tex
Commit for the sake of commit.
Update paper.tex

@@
 \begin{align*}
-\frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln n - n \ln n - cn)^2} = \frac{\sigma_X^2}{(n \ln n(\beta - 1) - cn)^2}
+\frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln n - n \ln n - O(n))^2} = \frac{\sigma_X^2}{(n \ln n(\beta - 1) - O(n))^2} \\
+&= \frac{\sigma_X^2}{n^2 (\ln n)^2 (\beta - 1)^2 - 2 O(n^2) \ln n (\beta - 1) + O(n^2)} \\
+&= \frac{\sigma_X^2}{n^2} \frac{1}{(\ln n)^2(\beta-1)^2-2O(1) \ln n (\beta-1) + O(1)} \\
+&= \frac{\sigma_X^2}{n^2} \frac{1}{(\ln n)^2(\beta-1)^2-(\beta-1)O(\ln n)} \\
+&\leq \frac{\pi^2}{(\beta-1)^2 O((\ln n)^2)-(\beta-1)O(\ln n)}
 \end{align*}
+
+\section*{Problem 4.14}
+
+We start by analyzing the root-to-leaf paths in the recursion for QuickSort. We want to show that a path from the root to a leaf is $O[lgn]$ with high probability.\\
+Looking on the path from the root to a specific leaf, we define $X_i$ to be 1 if the node on level $i$ splits the current array into 2 parts, where both of them are at most 3 times bigger than the other.
+Define $p_i$ to be the probability of $X_i$ being 1, which happens if the pivot was chosen in the middle half (between $\frac{1}{4}$ and $\frac{3}{4}$), so for every $i$ we have $p_i=\frac{1}{2}$.
+Define $X=\Sigma_{i=1}^kX_i$.
+So $X$ is the number of good splits in the first $k$ levels.
+We want to show that for $k=O(lgn)$ there is a high chance that we've reached a leaf.
+In the worst case, when we have $X_i=1$ we reduce the array to $\frac{3}{4}$, so when $X\geq lg_{\frac{3}{4}}n$ we know that we have reached a leaf.
+The probability that we haven't reached a leaf in the first $O(lgn)$ steps is thus (using chernoff bound) $Pr[X<lg_{\frac{3}{4}}n]=Pr[X<(1-\delta)O(lgn)]<exp(-O(lgn)\delta^2/2)=O(m^{-1})$.
+So with high n there is a high probability that we will have reached a leaf in $O(lgn)$ steps. So we can conclude that RandQS runs in time $O(nlgn)$.
 
@@
 
+\subsection*{Random Treaps}
+A treap is a special form of a binary search tree that is also a heap.\\
+When we need to store items with totally ordered keys, and later search for them, a binary search tree is a good option for a data structure.
+But there's many different ways to build a binary tree with a set of items.
+The more balanced a binary tree is, the better, because we need less time to search for item.
+Treaps are a randomized way of building a balanced binary tree.\\
+When building a treap, each item gets a random number as their priority, and then a tree is build, that is a binary search tree with respect to the keys and a heap with respect to the priorities.\\
+Proof that such a tree exists: When given a set of items with keys $k_i$ and priorities $p_i$, we can construct such treap: First we take the item with the highest priority.
+We call that item $x$.
+$x$ has to be at the root of the tree, because it has the highest priority.
+All items with key less than $k_x$ has to be in the left subtree, and all items with keys more than $k_x$ has to be in the right subtree.
+Now we can recursively build both the left subtree and the right subtree (the items to the left and right with the highest priorities has to be at the root of the subtrees, and we can divide the remaining items according to the keys).
+This proves that from any given set of items with keys and priorities, we can construct a treap containing those items, and furthermore it proves that it can be done in a unique way.\\
+With random priorities, treaps work very much like the Quicksort algorithm.
+Byt choosing random priorities to each of the items, it becomes random which of the items gets the highest priority.
+That item has to be in the root of the tree, and all other items has to be in either the left or right subtree according to whether their keys are less or more than the key of the item with the highest priority.
+This works in the same way as choosing a random item and partitioning according to it in the Quicksort algorithm.
+After the partitioning, we do the same recursively for both subtrees in the treap individually, like we do the same recursively for both partitions in the Quicksort algorithm.
+So treaps and the quicksort algorithm work in the same way.
+Since we know that the quicksort algorithm has expected $O(logn)$ layers, we can conclude that a random treap gets and expected height of $O(logn)$.\\
+\textbf{Lemma 8.6:} Let T be a random treap for a set S of size n. For an element $x\in S$ having rank k, $$E[depth(x)]=H_k+H_{n-k+1}-1$$
+$L_x$ = length of the left spine of the right sub-tree of x.\\
+$R_x$ = lenght of the right spine of the left sub-tree of x.\\
+\textbf{Lemma 8.7:} Let T be a random treap for a set S of size n. For an element $x\in S$ of rank k, $$E[R_x]=1-\frac{1}{k}$$ and $$E[L_x]=1-\frac{1}{n-k+1}.$$
+\textbf{Theorem 8.8:} Let T be a random treap for a set S of size n.
+\begin{enumerate}\item The expected time for a FIND, INSERT, or DELETE operation on T is $O(logn)$.\item The expected number of rotations required during an INSERT or DELETE operation is at most 2.\item The expected time for a JOIN, PASTE, or SPLIT operation involving sets $S_1$ and $S_2$ of sizes n and m, respectively, is $O(logn+logm)$.\end{enumerate}
+
 \end{document}

### 2017-05-22 assignments/3/paper.tex
Added problem 4

--- 2017-05-15 assignments/3/paper.tex
@@
 \begin{align*}
-\frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln n - n \ln n - cn)^2} = \frac{\sigma_X^2}{(n \ln n(\beta - 1) - cn)^2}
+\frac{\sigma_X^2}{(\beta n \ln (n) - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln (n) - n \ln(n) - O(n))^2} = \frac{\sigma_X^2}{(n \ln(n)(\beta - 1) - O(n))^2} \\
+&= \frac{\sigma_X^2}{n^2 \ln^2(n) (\beta - 1)^2 + O(n^2) - O(n^2 \ln(n))} \\
+&\leq \frac{\sigma_X^2}{O(n^2 \ln^2(n)) - O(n^2 \ln(n))} \\
+&= \frac{\sigma_X^2}{O(n^2 \ln^2(n)} \\
+&\leq \lim_{n \rightarrow \infty} \left( O \left( \frac{\sigma_X^2}{n^2} \right) \right) \frac{1}{O(\ln^2(n)} \\
+& = O \left( \frac{1}{\ln^2(n)} \right)
 \end{align*}
+The last two steps follow from $\sigma_X^2/n^2$ being monotonically increasing and converging for large enough $n$.
+
+\section*{Problem 3}
+Consider a treap $T=(e_1, e_2, ..., e_n) = ((v_1,p_1), (v_2,p_2), ..., (v_n,p_n))$ with an element $e_k = (v_k, -\infty) \in T$ of rank $k$.
+We wish to analyze the expected depth of $e_j$ in $T$.
+Let $X_i = 1$ if $e_i$ is an ancestor of $e_k$ and $X_i=0$ otherwise.
+Then the depth of $e_k$ is given by $X=1 + \sum_{i=1}^{n}X_i$ and
+\begin{align*}
+\mathbf{E}\left[ X \right] = 1 + \mathbf{E}\left[ \sum\limits_{i=1}^{n}X_i \right] = 1 + \sum\limits_{i=1}^{n}\mathbf{E}\left[X_i \right] 
+\end{align*}
+Since $e_k$ has rank $k$, $e_i$ is an ancestor of $e_k$ exactly when $e_i$ has the highest priority of the elements $e_i, ..., e_k$, when $i < k$ and $e_k, ..., e_i$ otherwise.
+Since the priorities are chosen at random, the probabilities of these events are given by $1/(k-i)$ and $1/(i-k)$, respectively.
+We then have $\mathbf{Pr}\left[ X_i = 1 \right] = 1 / |i-k|$ and
+\begin{align*}
+1 + \sum\limits_{i=1}^{n}\mathbf{E}\left[X_i \right] &= 1 + \sum\limits_{i=1}^{n}\mathbf{Pr}\left[X_i = 1 \right] \\
+&= 1 + \sum\limits_{i=1}^{n} 1 / |i-k| \\
+&= 1 +\sum\limits_{i=1}^{k-1} \frac{1}{k-i} + \sum\limits_{i=k+1}^{n} \frac{1}{i-k} \\
+&= 1 +H_{k-1} + H_{n-k}
+\end{align*}
+Subtracting the expected depth of $e_k$ from the expected depth of an element with rank $k$ and a random priority (lemma 8.6) yields
+\begin{align*}
+1 + H_{k-1} + H_{n-k} - H_k - H_{n-k+1} + 1 &= 1 + (H_{k-1} - H_k) + (H_{n-k} - H_{n-k+1}) + 1 \\
+&= 1 -\frac{1}{k} +1 - \frac{1}{n-k+1} \\
+&= \mathbf{E}\left[ R_x \right] + \mathbf{E}\left[ L_x \right]
+\end{align*}
+which exactly is the expected number of rotations required when deleting an element of rank $k$.
+This makes sense, because when deleting the element, we change its priority to $\infty$ and make rotations until it reaches the bottom.
+Each rotation moves it down 1 level.
+So the depth of the bottom (when the priority is negative infinity) should be the depth when we started plus the number of rotations.
+
+\section*{Problem 4}
+Consider the 2-level hashing scheme with $h_a$ used for the hash-function.
+Pick random, odd $a \in [2^w]$, and let $u = 2^w$ and $t = 2^\ell$ where $\ell \leq w$.
+Let $C_{x,y} = h_a(x) == h_a(y)$, i.e.
+if there is a collision between $x$ and $y$ and $C = \sum_{x \neq y} C_{x,y}$ denote the total number of collisions.
+Then $\mathbf{E} \left[ C \right] = \mathbf{E}\left[ \sum_{x \neq y} C_{x,y} \right] \leq \sum_{x \neq y} 2/2^{\ell} = \binom{n}{2}2/2^\ell = n(n-1)/2^\ell$.
+For a set of size $n$, we choose $\ell = \lfloor \log_2(n-1) \rfloor + 2$, for which $\mathbf{E}[C] = n(n-1)/2^{\lfloor \log_2(n-1) \rfloor + 2} \leq n(n-1)/2^{\log_2(n-1) + 1} = n/2$.
+Pick $h_a$ s.t.
+$C \leq n$.
+Then from applying Markov's inequality we get that $\mathbf{Pr}[X \geq n] \leq \frac{n/2}{n} = 1/2$, so we expect at most 2 tries before $C \leq n$.
+Letting $t_i = n_i(n_i-1)$ where $n_i$ is the number of collisions in bucket $i$.
+The amount of space required by this method is given by $\Sigma_{i \in [2^\ell]}t_i = \Sigma_{i \in [2^\ell]} 2 \binom{n_i}{2} = 2 \Sigma_{i \in [2^\ell]} \binom{n_i}{2} = 2C \leq 2n = O(n)$.
 
@@
 
+\subsection*{Hashing}
+Consider $X \subseteq [u]$, $|X| = n$ keys drawn from universe $[u] = [1, \hdots, u-1]$.
+We wish to store $X$ so as to decide $x \in X$ by using a hash-function $h : [u] \rightarrow [t]$.
+We say $h$ is universal if given $\forall x,y \in [u]$ and $x \neq y$ then $\mathbf{Pr}\left[ h(x) = h(y) \right] \leq 1/t$.
+The following are examples of hash functions:
+\begin{enumerate}
+\item Pick a prime $p \geq u$ and $a \in [p]_+ = [1, \hdots, p-1]$, $b \in [u]$ uniformly at random. Then $h_{a,b}(x) = ((ax + b) \mod p) \mod t$, then $h_{a,b}$ is universal.
+\item Let $u=2^w$, $t=2^\ell$ and $\ell \leq w$. Pick a random odd $a \in [2^w]$. Then $h_a(x) = (a * x) \gg (w - \ell)$. Then $\mathbf{Pr}\left[ h_a(x) = h_a(y) \right] \leq 2/2^\ell$.
+\end{enumerate}
+\paragraph{Naive hashing} We want a worst-case query time of $O(1)$ for deciding the membership problem.
+Given a hash-function $h$, we construct an array $T[t]$, and set $T[h(x)] = x$ for each $x \in X$.
+Then $x \in X \Leftrightarrow T[h(x)] = x$.
+We pick $t$ large enough to expect no collisions.
+Let $C_{x,y} = h(x) == h(y)$, then the total number of collisions is denoted by $C = \sum_{x \neq y} C_{x,y}$.
+Now $\mathbf{E}\left[ C \right] = \mathbf{E}\left[ \sum_{x \neq y} C_{x,y} \right] \leq \sum_{x \neq y} 1/t = \binom{n}{2}/t$.
+By picking $t=n(n-1)=2\binom{n}{2}$, we get $\mathbf{Pr} \left[ C \geq 1 \right] \leq \mathbf{E}\left[ C \right] / 1 \leq 1/2$.
+If $C \geq 1$, then there is some collision and we try with a new $h$.
+This approach uses $O(n^2)$ space.
+
+\paragraph{Two-level hashing} Let $n=t$ and $h$ be a hash function. We construct a hashmap as follows:
+
+\begin{algorithm}[H]
+ \For{each element $i \in [t]$}{
+  store $B_i = \left\{ x \in X \; | \; h(x) = i \right\}$
+  }
+ \caption{Two-level hashing algorithm}
+\end{algorithm}
+To answer queries of the type $y \in X$, we can check $y \in B_{h(y)}$.
+Note that $\mathbf{E} \left[ |B_{h(y)}| \right] = n/t$.
+Pick $t = n-1$.
+Let $C$ be defined as before.
+Then $\mathbf{E} \left[ C \right] = \binom{n}{2}/t = n/2$.
+If $C > n$, try new $h$.
+Now $C \leq n$.
+For $t_i=n_i(n_i-1) = 2\binom{n_i}{2}$, use array $T_i[t_i]$ for each bucket.
+Now determining whether $y \in X \Leftrightarrow T_i[h_i(y)] = T[h(y)][h_i(y)]$, assuming $h_i$ has no collisions.
+Otherwise choose a new $h_i$.
+The space required by this approach is
+\begin{align*}
+\sum_{i \in [t]} t_i = \sum_{i \in [t]}2\binom{n_i}{2} = 2\sum_{i \in [t]}\binom{n_i}{2} = 2C \leq 2n
+\end{align*}
+which is linear in $n$.
+
 \end{document}

### 2017-05-22 assignments/3/paper.tex

@@
 
-\section*{Exercise 3.10}
-We wish to find an upper bound on $\mathbf{Pr} \left[ X > \beta n \ln n \right] \leq \mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \beta n \ln n - \mu_X \right]$.
+\section*{1)}
+We wish to determine an upper bound on $\mathbf{Pr} \left[ X > \beta n \ln n \right] \leq \mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \beta n \ln n - \mu_X \right]$.
 Now, setting $t\sigma_X = \beta n \ln n - \mu_X$, we get $t = \frac{\beta n \ln n - \mu_X}{\sigma_X}$.
@@
 
-\section*{Problem 3}
+\section*{2)}
+
+We start by analyzing the root-to-leaf paths in the recursion for QuickSort.
+We want to show that a path from the root to a leaf is $O(\lg n)$ with high probability.
+Looking at the path from the root to a specific leaf, we define $X_i$ to be 1 if the node on level $i$ splits the current array into 2 parts, where one of them are at most 3 times bigger than the other.
+Define $p_i$ to be the probability of $X_i$ being 1, which happens if the pivot was chosen in the middle half (between $\frac{1}{4}$ and $\frac{3}{4}$), so for every $i$ we have $p_i=\frac{1}{2}$.
+Define $X=\Sigma_{i=1}^k X_i$.
+So $X$ is the number of good splits in the first $k$ levels.
+We want to show that for $k=O(\lg n)$ there is a high chance that we have reached a leaf.
+In the worst case, when we have $X_i=1$ we reduce the array to $\frac{3}{4}$, so when $X\geq lg_{\frac{3}{4}}n$ we know that we have reached a leaf.
+The probability that we have not reached a leaf in the first $O(\lg n)$ steps is thus (by using Chernoff Bound) $\mathbf{Pr}[X < \log_{\frac{3}{4}}n]=\mathbf{Pr}[X<(1-\delta)O(\lg n)] < exp(-O(\lg n)\delta^2/2)=O(m^{-1})$.
+That is, for large $n$ there is a high probability that we will have reached a leaf in $O(\lg n)$ steps.
+So we can conclude that RandQS runs in time $O(n \lg n)$.
+
+\section*{3)}
 Consider a treap $T=(e_1, e_2, ..., e_n) = ((v_1,p_1), (v_2,p_2), ..., (v_n,p_n))$ with an element $e_k = (v_k, -\infty) \in T$ of rank $k$.
@@
 
-\section*{Problem 4}
-Consider the 2-level hashing scheme with $h_a$ used for the hash-function.
-Pick random, odd $a \in [2^w]$, and let $u = 2^w$ and $t = 2^\ell$ where $\ell \leq w$.
+\section*{4)}
+Consider the 2-level hashing scheme with $h_a$ used for the hash-function. Pick random, odd $a \in [2^w]$, and let $u = 2^w$ and $t = 2^\ell$ where $\ell \leq w$.
+
+\begin{algorithm}[H]
+ \For{each element $i \in [t]$}{
+  store $B_i = \left\{ x \in X \; | \; h(x) = i \right\}$
+  }
+ \caption{Two-level hashing algorithm}
+\end{algorithm}
+
+Each bucket $B_i$ is internally represented by a secondary hash-table, $T_i[t_i]$.
 Let $C_{x,y} = h_a(x) == h_a(y)$, i.e.
-if there is a collision between $x$ and $y$ and $C = \sum_{x \neq y} C_{x,y}$ denote the total number of collisions.
+if there is a collision between $x$ and $y$ and $C = \sum_{x \neq y} C_{x,y}$ denote t he total number of collisions.
 Then $\mathbf{E} \left[ C \right] = \mathbf{E}\left[ \sum_{x \neq y} C_{x,y} \right] \leq \sum_{x \neq y} 2/2^{\ell} = \binom{n}{2}2/2^\ell = n(n-1)/2^\ell$.
@@
 Then from applying Markov's inequality we get that $\mathbf{Pr}[X \geq n] \leq \frac{n/2}{n} = 1/2$, so we expect at most 2 tries before $C \leq n$.
-Letting $t_i = n_i(n_i-1)$ where $n_i$ is the number of collisions in bucket $i$.
-The amount of space required by this method is given by $\Sigma_{i \in [2^\ell]}t_i = \Sigma_{i \in [2^\ell]} 2 \binom{n_i}{2} = 2 \Sigma_{i \in [2^\ell]} \binom{n_i}{2} = 2C \leq 2n = O(n)$.
+Letting $t_i = 2^{\lfloor \log_2 n_i(n_i-1) \rfloor} \leq 2^{\log_2 n_i(n_i-1)+1} = 2n_i(n_i-1) = 4 \binom{n_i}{2}$ where $n_i$ is the number of collisions in bucket $i$.
+Another hash function $h_{a_i}$ is chosen so that it perfectly hashes the elements of bucket $i$ into the hash-table $T_i$.
+The amount of space required by this method is given by $\Sigma_{i \in [2^\ell]}t_i = \Sigma_{i \in [2^\ell]} 4 \binom{n_i}{2} = 4 \Sigma_{i \in [2^\ell]} \binom{n_i}{2} = 4C \leq 4n = O(n)$.
 
@@
 
+\subsection*{Random Treaps}
+A treap is a special form of a binary search tree that is also a heap.\\
+When we need to store items with totally ordered keys, and later search for them, a binary search tree is a good option for a data structure.
+But there's many different ways to build a binary tree with a set of items.
+The more balanced a binary tree is, the better, because we need less time to search for item.
+Treaps are a randomized way of building a balanced binary tree.\\
+When building a treap, each item gets a random number as their priority, and then a tree is build, that is a binary search tree with respect to the keys and a heap with respect to the priorities.\\
+Proof that such a tree exists: When given a set of items with keys $k_i$ and priorities $p_i$, we can construct such treap: First we take the item with the highest priority.
+We call that item $x$.
+$x$ has to be at the root of the tree, because it has the highest priority.
+All items with key less than $k_x$ has to be in the left subtree, and all items with keys more than $k_x$ has to be in the right subtree.
+Now we can recursively build both the left subtree and the right subtree (the items to the left and right with the highest priorities has to be at the root of the subtrees, and we can divide the remaining items according to the keys).
+This proves that from any given set of items with keys and priorities, we can construct a treap containing those items, and furthermore it proves that it can be done in a unique way.\\
+With random priorities, treaps work very much like the Quicksort algorithm.
+Byt choosing random priorities to each of the items, it becomes random which of the items gets the highest priority.
+That item has to be in the root of the tree, and all other items has to be in either the left or right subtree according to whether their keys are less or more than the key of the item with the highest priority.
+This works in the same way as choosing a random item and partitioning according to it in the Quicksort algorithm.
+After the partitioning, we do the same recursively for both subtrees in the treap individually, like we do the same recursively for both partitions in the Quicksort algorithm.
+So treaps and the quicksort algorithm work in the same way.
+Since we know that the quicksort algorithm has expected $O(logn)$ layers, we can conclude that a random treap gets and expected height of $O(logn)$.\\
+\textbf{Lemma 8.6:} Let T be a random treap for a set S of size n. For an element $x\in S$ having rank k, $$E[depth(x)]=H_k+H_{n-k+1}-1$$
+$L_x$ = length of the left spine of the right sub-tree of x.\\
+$R_x$ = lenght of the right spine of the left sub-tree of x.\\
+\textbf{Lemma 8.7:} Let T be a random treap for a set S of size n. For an element $x\in S$ of rank k, $$E[R_x]=1-\frac{1}{k}$$ and $$E[L_x]=1-\frac{1}{n-k+1}.$$
+\textbf{Theorem 8.8:} Let T be a random treap for a set S of size n.
+\begin{enumerate}\item The expected time for a FIND, INSERT, or DELETE operation on T is $O(logn)$.\item The expected number of rotations required during an INSERT or DELETE operation is at most 2.\item The expected time for a JOIN, PASTE, or SPLIT operation involving sets $S_1$ and $S_2$ of sizes n and m, respectively, is $O(logn+logm)$.
+\end{enumerate}
+
 \end{document}

### 2017-05-24 assignments/4/paper.tex
Assignment 4 stub.

--- 2017-05-22 assignments/3/paper.tex
@@
 \title{Randomized Algorithms \\
-       Assignment 3}
+       Assignment 4}
 
@@
 \begin{abstract}
-  This is the third weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
+  This is the fourth weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
 \end{abstract}
@@
 
-\section*{Exercise 3.10}
-We wish to find an upper bound on $\mathbf{Pr} \left[ X > \beta n \ln n \right] \leq \mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \beta n \ln n - \mu_X \right]$.
-Now, setting $t\sigma_X = \beta n \ln n - \mu_X$, we get $t = \frac{\beta n \ln n - \mu_X}{\sigma_X}$.
-Inserting and applying Chebyshev's inequality yields
-\begin{align*}
-\mathbf{Pr} \left[ \left| X - \mu_X \right| \geq \frac{\beta n \ln n - \mu_X}{\sigma_X} \sigma_X \right] \leq \frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2}
-\end{align*}
-Inserting $\mu_X = \mathbf{E} \left[ X \right] = n \ln n + O(n)$ gives
-\begin{align*}
-\frac{\sigma_X^2}{(\beta n \ln n - \mu_X)^2} &= \frac{\sigma_X^2}{(\beta n \ln n - n \ln n - O(n))^2} = \frac{\sigma_X^2}{(n \ln n(\beta - 1) - O(n))^2} \\
-&= \frac{\sigma_X^2}{n^2 (\ln n)^2 (\beta - 1)^2 - 2 O(n^2) \ln n (\beta - 1) + O(n^2)} \\
-&= \frac{\sigma_X^2}{n^2} \frac{1}{(\ln n)^2(\beta-1)^2-2O(1) \ln n (\beta-1) + O(1)} \\
-&= \frac{\sigma_X^2}{n^2} \frac{1}{(\ln n)^2(\beta-1)^2-(\beta-1)O(\ln n)} \\
-&\leq \frac{\pi^2}{(\beta-1)^2 O((\ln n)^2)-(\beta-1)O(\ln n)}
-\end{align*}
+\section*{1)}
 
-\section*{Problem 4.14}
+\section*{2)}
 
-We start by analyzing the root-to-leaf paths in the recursion for QuickSort. We want to show that a path from the root to a leaf is $O[lgn]$ with high probability.\\
-Looking on the path from the root to a specific leaf, we define $X_i$ to be 1 if the node on level $i$ splits the current array into 2 parts, where both of them are at most 3 times bigger than the other.
-Define $p_i$ to be the probability of $X_i$ being 1, which happens if the pivot was chosen in the middle half (between $\frac{1}{4}$ and $\frac{3}{4}$), so for every $i$ we have $p_i=\frac{1}{2}$.
-Define $X=\Sigma_{i=1}^kX_i$.
-So $X$ is the number of good splits in the first $k$ levels.
-We want to show that for $k=O(lgn)$ there is a high chance that we've reached a leaf.
-In the worst case, when we have $X_i=1$ we reduce the array to $\frac{3}{4}$, so when $X\geq lg_{\frac{3}{4}}n$ we know that we have reached a leaf.
-The probability that we haven't reached a leaf in the first $O(lgn)$ steps is thus (using chernoff bound) $Pr[X<lg_{\frac{3}{4}}n]=Pr[X<(1-\delta)O(lgn)]<exp(-O(lgn)\delta^2/2)=O(m^{-1})$.
-So with high n there is a high probability that we will have reached a leaf in $O(lgn)$ steps. So we can conclude that RandQS runs in time $O(nlgn)$.
-
-\pagebreak
-
-\section*{Summary}
-\subsection*{Coupon Collector's Problem}
-Given $n$ types of coupons, a trial means picking a coupon independently and at random.
-Let $X = \text{\# of trials to get all } n \text{ coupon types}$.
-For $i=0, \hdots, n-1$, the $i$th epoch consists of the trials starting just after the $i$th success and ending in the $(i+1)$th success.
-Define $X_i = \text{\# of trials in the } i\text{th epoch.}$, then $X=\sum_{i=0}^{n-1} X_i$.
-Now
-\begin{align*}
-\mathbf{E}\left[ X \right] = \sum_{i=0}^{n-1} \mathbf{E}\left[ X_i \right] = \sum_{i=0}^{n-1} \frac{n}{n-i} = \sum_{i=1}^{n} \frac{n}{i} =  n \sum_{i=1}^{n} \frac{1}{i} = n H_{n} = n \ln n + O(n)
-\end{align*}
-Let $\varepsilon_i^r$ denote the event that coupon type $i$ is not picked in the first $r$ trials. Then
-\begin{align*}
-\mathbf{Pr}\left[ \varepsilon_i^r \right] = \left(1-\frac{1}{n}\right)^r \leq e^{-r/n}
-\end{align*}
-Pick $r=\beta n \ln n$, where $\beta$ is a constant, then $\mathbf{Pr}\left[ \varepsilon_i^r \right] \leq e^{-\beta \ln n} = n^{-\beta}$ and
-\begin{align*}
-\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq \sum_{i=1}^n \mathbf{Pr}\left[ \varepsilon_i^r \right] \leq n^{1-\beta}
-\end{align*}
-For $\beta=2$, we get $\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq 1/n$.
-
-\subsection*{Chernoff Bounds}
-Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$.
-Let $X=\sum_{i=1}^n X_i$ and $\mu = \mathbf{E} \left[ X \right] = \sum_{i=1}^n p_i$.
-Then
-\begin{align*}
-\mathbf{Pr}\left[ X > (1-\delta)\mu_X \right]
-&= \mathbf{Pr}\left[ e^{tX} > e^{t(1-\delta)\mu_X} \right]
-< \frac{\mathbf{E}\left[ e^{tX} \right]}{e^{t(1-\delta)\mu_X}}
-= \frac{\prod_{i=1}^n \mathbf{E}\left[ e^{tX_i} \right]}{e^{t(1-\delta)\mu_X}} \\
-&= \frac{\prod_{i=1}^n \left( p_i e^t + 1-p_i \right)}{e^{t(1-\delta)\mu_X}}
-= \frac{\prod_{i=1}^n \left( 1 + p_i (e^t - 1) \right)}{e^{t(1-\delta)\mu_X}} \\
-&\leq \frac{\prod_{i=1}^n e^{p_i (e^t-1)}}{e^{t(1-\delta)\mu_X}}
-\leq \left( \frac{e^{(e^t-1)}}{e^{t(1-\delta)}} \right)^{\mu_X}
-= \left( \frac{e^{\delta}}{\left( 1 + \delta \right)^{1+\delta}} \right)^{\mu_X}
-\end{align*}
-Furthermore
-\begin{align*}
-\mathbf{Pr}\left[ X < (1-\delta)\mu_X \right]
-&= \mathbf{Pr}\left[ -X < -(1-\delta)\mu_X \right]
-= \mathbf{Pr}\left[ e^{-tX} < e^{-t(1-\delta)\mu_X} \right] \\
-&< \left( \frac{e^{-\delta}}{\left( 1-\delta \right)^{1-\delta}} \right)^\mu
-= \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right) \ln (1-\delta)}} \right)^\mu \\
-&< \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right)\left( -\delta - \frac{\delta^2}{2} \right)}} \right)^\mu
-=  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2 + \frac{\delta^3}{2}}  } \right)^\mu \\
-&<  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2}  } \right)^\mu = \left( \frac{1}{e^{-\frac{\delta^2}{2} + \delta^2}  } \right)^\mu
-= \left( e^{\frac{\delta^2}{2} - \delta^2} \right)^\mu \\
-&= e^{\mu\frac{\delta^2}{2} - \mu\delta^2} = e^{-\mu\sigma^2/2}
-\end{align*}
-
-\subsection*{Set-Balancing}
-Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t.
-$|| A\overline{b} ||_{\infty}$ is minimized.
-Pick each entry of $\overline{b}$ to be $1$ with probability $1/2$ and $-1$ otherwise.
-Now, $\mathbf{E} \left[ || A\overline{b} ||_{\infty} \right] = 0$.
-Consider a row $i$, $A_{i}$, of $A$, and assume for now that all entries are 1.
-Let $X$ be the \# of 1's in $\overline{b}$.
-Now
-\begin{align*}
-\mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] &\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \; \lor \; X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
-&\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \right] + \mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
-&\leq \frac{1}{n^2} + \frac{1}{n^2} = \frac{2}{n^2}
-\end{align*}
-and the third inequality follows from
-\begin{align*}
-\mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right]
-&= \mathbf{Pr} \left[ X < \left( 1 - \frac{2\sqrt{2n \ln n}}{n} \right)\frac{n}{2} \right]
-< e^{-\frac{n}{2} \left( \frac{2\sqrt{2n \ln n}}{n}\right)^2\frac{1}{2}} \\
-&< e^{-\frac{n}{4} \left( \frac{8n \ln n}{n^2}\right)} = e^{-2\ln n} = 1/n^2
-\end{align*}
-where the first inequality follows from applying the second Chernoff bound to the inequality with $\delta=\frac{2\sqrt{2n \ln n}}{n}$ and $\mu=n/2$.
-Finally applying an union bound over all rows we get
-\begin{align*}
-\mathbf{Pr} \left[ || A\overline{b} ||_{\infty} > 2\sqrt{2n \ln n} \right] \geq \sum_{i=1}^n \mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] \leq \sum_{i=1}^n \frac{2}{n^2} = \frac{2}{n}
-\end{align*}
-
-\subsection*{Random Treaps}
-A treap is a special form of a binary search tree that is also a heap.\\
-When we need to store items with totally ordered keys, and later search for them, a binary search tree is a good option for a data structure.
-But there's many different ways to build a binary tree with a set of items.
-The more balanced a binary tree is, the better, because we need less time to search for item.
-Treaps are a randomized way of building a balanced binary tree.\\
-When building a treap, each item gets a random number as their priority, and then a tree is build, that is a binary search tree with respect to the keys and a heap with respect to the priorities.\\
-Proof that such a tree exists: When given a set of items with keys $k_i$ and priorities $p_i$, we can construct such treap: First we take the item with the highest priority.
-We call that item $x$.
-$x$ has to be at the root of the tree, because it has the highest priority.
-All items with key less than $k_x$ has to be in the left subtree, and all items with keys more than $k_x$ has to be in the right subtree.
-Now we can recursively build both the left subtree and the right subtree (the items to the left and right with the highest priorities has to be at the root of the subtrees, and we can divide the remaining items according to the keys).
-This proves that from any given set of items with keys and priorities, we can construct a treap containing those items, and furthermore it proves that it can be done in a unique way.\\
-With random priorities, treaps work very much like the Quicksort algorithm.
-Byt choosing random priorities to each of the items, it becomes random which of the items gets the highest priority.
-That item has to be in the root of the tree, and all other items has to be in either the left or right subtree according to whether their keys are less or more than the key of the item with the highest priority.
-This works in the same way as choosing a random item and partitioning according to it in the Quicksort algorithm.
-After the partitioning, we do the same recursively for both subtrees in the treap individually, like we do the same recursively for both partitions in the Quicksort algorithm.
-So treaps and the quicksort algorithm work in the same way.
-Since we know that the quicksort algorithm has expected $O(logn)$ layers, we can conclude that a random treap gets and expected height of $O(logn)$.\\
-\textbf{Lemma 8.6:} Let T be a random treap for a set S of size n. For an element $x\in S$ having rank k, $$E[depth(x)]=H_k+H_{n-k+1}-1$$
-$L_x$ = length of the left spine of the right sub-tree of x.\\
-$R_x$ = lenght of the right spine of the left sub-tree of x.\\
-\textbf{Lemma 8.7:} Let T be a random treap for a set S of size n. For an element $x\in S$ of rank k, $$E[R_x]=1-\frac{1}{k}$$ and $$E[L_x]=1-\frac{1}{n-k+1}.$$
-\textbf{Theorem 8.8:} Let T be a random treap for a set S of size n.
-\begin{enumerate}\item The expected time for a FIND, INSERT, or DELETE operation on T is $O(logn)$.\item The expected number of rotations required during an INSERT or DELETE operation is at most 2.\item The expected time for a JOIN, PASTE, or SPLIT operation involving sets $S_1$ and $S_2$ of sizes n and m, respectively, is $O(logn+logm)$.\end{enumerate}
+\section*{3)}
 

### 2017-09-07 assignments/4/paper.tex
Added 5 + 6

@@
 \section*{2)}
+Let $||y||$ denote the rounding of $y$ to the the nearest integer.
+Using this scheme will produce a feasible solution to the ILP since $||x_{i0}|| + ||x_{i1}|| = 1$ for all $i$.
+For some instance of the wiring problem, let $w_0$, $\hat{w}$ and $\bar{w}$ denote the optimal value of the ILP, LP-relaxation and rounded solution respectively and let $b$ be any boundary having $w_0$ wires crossing.
+Furthermore let $T_b=T_{b0} \cup T_{b1}$ and since $T_{b0} \cap T_{b1} = \emptyset$, we have
+\begin{align*}
+\bar{w} &= \sum_{i \in T_{b0}} || \hat{x}_{i0} || + \sum_{i \in T_{b1}} || \hat{x}_{i1} || \\
+&=\sum_{i \in T_{b}} \left( ||\hat{x}_{i0} || + ||\hat{x}_{i1}|| \right) - \sum_{i \in T_{b0}} ||\hat{x}_{i1} || - \sum_{i \in T_{b1}} ||\hat{x}_{i0} || \\
+&\leq\sum_{i \in T_{b}} \left( x_{i0} + x_{i1} \right) \\
+&=\sum_{i \in T_{b0}} x_{i0} + \sum_{i \in T_{b1}} x_{i0} + \sum_{i \in T_{b0}} x_{i1} + \sum_{i \in T_{b1}} x_{i1} \\
+&=w_0 + \sum_{i \in T_{b0}} x_{i1} + \sum_{i \in T_{b1}} x_{i0} \\
+&=w_0 + \sum_{i \in T_{b0}} x_{i0} + \sum_{i \in T_{b1}} x_{i1} - \sum_{i \in T_{b0}} 1 - \sum_{i \in T_{b1}} 1 \\
+&\leq 2w_0
+\end{align*}
 
 \section*{3)}
+The following ILP produces a solution to a set-cover instance encoded by $\mathbf{M}$:
+\begin{align*}
+\text{minimize} \;& ||\mathbf{c}||_1 \\
+\text{s.t.} \;& \mathbf{M}_{i,:} \cdot \mathbf{c} \geq 1 \; \; \; \forall i \in \left\{ 1 \hdots n \right\} \\
+\text{where} \;& \mathbf{c}_i \in \left\{ 0,1 \right\}
+\end{align*}
+A randomized algorithm for finding a set-cover first solves the $LP$ relaxation of the above problem, by relaxing the constraint that on $\mathbf{c}$ to the domain of reals and outputting a rounded solution $\overline{\mathbf{c}} \in \left\{ 0,1 \right\}^m$ where $\mathbf{Pr}\left[\overline{\mathbf{c}}_i = 1 \right] = \mathbf{c}_i$.
+
+Let $L_i = \mathbf{M}_{i,:} \cdot \overline{ \mathbf{c} }$.
+If we can bound the probability that a particular set is not covered, we can apply union bound to bound the probability that not all sets are covered from below.
+The probability that $S_i$ is not covered is given by $\mathbf{Pr}\left[ L_i < 1 \right]$.
+Since the LHS is the sum of independent Poisson trials, we can apply Chernoff bound.
+Bounding $\mu_{L_i}$ from above, we get
+\begin{align*}
+\mu_{L_i} &= \mathbf{E}\left[ \mathbf{M}_{i,:} \cdot \overline{\mathbf{c}} \right] = \mathbf{E}\left[ \sum_{j=1}^{m} \mathbf{M}_{i,j} \cdot \overline{\mathbf{c}}_j \right] = \sum_{j=1}^{m} \mathbf{M}_{i,j} \cdot \mathbf{E}\left[ \overline{\mathbf{c}}_j \right] \\
+&= \sum_{j=1}^{m} \mathbf{M}_{i,j} \cdot \mathbf{Pr}\left[ \overline{\mathbf{c}}_j = 1 \right] = \sum_{j=1}^{m} \mathbf{M}_{i,j} \cdot \mathbf{c}_j \\
+&\leq || \mathbf{c} ||_1
+\end{align*} 
+Solving $1 = (1-\delta)\mu_{L_i}$ for $\mu_{L_i}$, we get $\delta = 1-1/\mu_{L_i}$. Applying Chernoff bound
+\begin{align*}
+\mathbf{Pr}\left[ L_i < 1 \right] &\leq \exp \left[ -\frac{1}{2}||\mathbf{c}_1|| \left( 1-\frac{1}{||\mathbf{c}_1||} \right)^2 \right] \\
+&= \exp \left[ -\frac{1}{2}||\mathbf{c}_1|| \left( 1+\frac{1}{||\mathbf{c}_1||^2} - \frac{2}{||\mathbf{c}_1||} \right) \right] \\
+&= \exp \left[ -\frac{||\mathbf{c}_1||}{2}-\frac{1}{2||\mathbf{c}_1||} + 1 \right]  \\
+&= \exp \left[ -\frac{1}{2||\mathbf{c}_1||} + \frac{1}{2} \right]  \\
+\end{align*}
+Meaning that $\mathbf{Pr}\left[ L_i \geq 1 \right] > 1-e^{-m^2}$. Finally, a union bound over all $S_i$ yields
+\begin{align*}
+\mathbf{Pr}\left[ \bigcup_{i=1}^{n} L_i \geq 1 \right] \leq n (1-e^{-m^2})
+\end{align*}
 

### 2017-09-07 assignments/5/paper.tex
Added 5 + 6

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[]{algorithm2e}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{alg}[thm]{Algorithm}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Randomized Algorithms \\
       Assignment 5}

\author{Viktor Hansen \& Simon Rueskov Schleicher}

\maketitle

\begin{abstract}
  This is the fifth weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{5.3)}
\paragraph{a)} Consider a graph $G$ with $n$ vertices and $nd/2$. Delete each vertex and its incident edges with probability $1-1/d$. Let $N_i=1$ if vertex $i$ is deleted and $N=\sum_{i=1}^{n} N_i$ be the total number of vertices after all deletions. Then
\begin{align*}
\mathbf{E}[N] = \sum_{i=1}^{n} \mathbf{E}[N_i] = \sum_{i=1}^{n} 1/d = n/d
\end{align*}
In a similar fashion, let $E_i=1$ if edge is is not deleted. Then
\begin{align*}
\mathbf{E}[E] &= \sum_{i=1}^{nd/2} \mathbf{E}[E_i] = \sum_{i=1}^{nd/2} 1-\mathbf{Pr}\left[ \text{edge } i \text{ belonging to a deleted node.} \right] \\
&= \sum_{i=1}^{nd/2} 1-(1-1/d^2) \\
&= \frac{n}{2d}
\end{align*}

\paragraph{b)} For any graph $G=(V,E)$, a trivial lower bound on the size of any independent set is $\max\left\{0,|V| - |E|\right\}$, since each edge can reduce the size of such an independent set by at most 1. Let $X$ and $Y$ denote the number of vertices and edges in $G$ after the deletion in a). Now, repeatedly pick a vertex, $v$, of $G$ and remove all its neighbours in $G$ remain and add $v$  to an initially empty set, $S$, until no vertices remain in $G$. Now $S$ is clearly an independent set, and its size is bounded from below by $X-Y$. Now
\begin{align*}
\mathbf{E}\left[ S \right] &\geq \mathbf{E}\left[ \max\left\{0,X-Y\right\} \right] = \mathbf{E}\left[ X \right] - \mathbf{E}\left[ Y \right] = \frac{n}{d} - \frac{n}{2d} = \frac{n}{2d}
\end{align*}
Since $\mathbf{E}\left[ S \right] \geq n/2d$, $S$ achieves a value of at least $n/2d$ with positive probability, by (2) of the probabilistic method. By (1) of the probabilistic method, this implies that an independent set of size at least $n/2d$ must exist.

\section*{2) Show that lemma 5.11 implies corollary 5.12}
Let $x_i = \frac{1}{d+1}$ for all $1 \leq i \leq n$. Then
\begin{align*}
x_i\prod_{(i,j) \in E} (1-x_j) &= \frac{1}{d+1} \left( 1-\frac{1}{d+1} \right)^{\deg(i)} \\
&\geq \frac{1}{d+1} \left( 1-\frac{1}{d+1} \right)^{d} \\
&\geq \frac{1}{d+1} \cdot \frac{1}{e} \geq p \geq  \mathbf{Pr} \left[ \varepsilon_i \right]
\end{align*}
By Lovasz Local Lemma,
\begin{align*}
\mathbf{Pr} \left[ \cap_{i=1}^{n} \overline{\varepsilon}_i  \right] \geq \prod_{i=1}^n \left( 1 - \frac{1}{d+1} \right) \geq \left( 1 - \frac{1}{d+1} \right)^n > 0
\end{align*}
when $\frac{1}{(d+1)e} \geq p \Leftrightarrow 1 \geq ep(d+1)$.

\section*{3) 5.13}
Using an approach similar to that of section 5.6, we adopt a computation-tree view of the algorithm. Each level of the tree corresponds to assigning a vertex to $A$ or $B$. If for any step, the algorithm can assign a vertex to $A$ or $B$ s.t. $\mathbf{Pr}\left[\text{\# edges crossing} < m/2\right] < 1$, the algorithm must terminate with $\text{\# edges crossing} \geq m/2$ at a leaf node. Theorem 5.1 tells us that this is the case at the root, so if we can maintain this invariant for each step, it is guaranteed to find a solution deterministically. Let $C_i, A_i, B_i$ and $U_i$ denote the size of the cut, $A$, $B$ and the set of unassigned vertices at each step $i$, for $1 \leq i \leq n$. Then $\mathbf{E}\left[ C_i \right] = \frac{1}{2} \mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right] + \frac{1}{2} \mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right]$ by employing the randomized partitioning strategy. It follows that $\mathbf{E}\left[ C_i \right] \leq \max \left\{ \mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{b+1} \right], \mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right] \right\}$ and thus choosing the assignment maximising the expectation can only decrease the probability of failure. The conditional expectation $\mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right]$ can be computed as follows:
\begin{align*}
\mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right] &= \left| A_{i+1} \times B_{i+1} \right| + \frac{1}{2} \left| U_{i+1}^2 \right| + \frac{1}{2} \left| A_{i+1} \times U_{i+1} \right|  + \frac{1}{2} \left| U_{i+1} \times B_{i+1} \right|\\
&= \left| A_{i+1} \times B_{i+1} \right| + \frac{1}{2} |\overline{A_{i+1} \times B_{i+1}}|
\end{align*}
and similarly for $\mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right]$. Since $\mathbf{E}\left[ C_i \right] \leq \mathbf{E}\left[ C_{i+1} \right]$ for all $1 \leq i \leq n$, we conclude that the the derandomised algorithm must produce a cut of size at least $m/2$.
\pagebreak

\section*{Summary May 18th and 23rd}
\subsection*{The Probabilistic Method}
\begin{enumerate}
\item If an object drawn from a universe has some property P with probability $> 0$, then some object in the universe has property P.
\item Any random var. achieves a value of at least (resp. at most) its expected value with positive probability.
\end{enumerate}

\subsection*{Max-Cut}
Given an undirected graph $G=(V,E)$, find a cut $V_1 \cup V_2$ of $V$ s.t. $\left\{ (u,v) \in E \; | \; u \in V_2, v \in V_2 \right\}$ is maximized. \\

\begin{alg}
For each $v \in V$, add it to $V_1$ w. prob. $1/2$, and add it to $V_2$ otherwise. \\
\end{alg}

\begin{thm}
$G$ has a cut with at least $m/2$ edges, where $m=|E|$.
\paragraph{Proof:} $\mathbf{E}[\text{\# edges crossing }(V_1,V_2)\text{ with above algo.}]=m/2$. Then by TPM, there exists a cut of $G$ with $\geq m/2$ edges crossing it.
\end{thm}

\subsection*{Max-SAT}
Given boolean formula in CNF with $m$ clauses, e.g. $(x_1 \lor \neg x_2 \lor x_3) \land (x_4 \lor x_1) \land \hdots$, and assume that no clause contains more than one occurrence of each variable. Then \\
\begin{thm}
Any B. formula in CNF with $m$ clauses each with at least $k$ literals has an assignment that satisfies at least $m(1-2^k)$ clauses. \\
\end{thm}

\begin{alg}
Set each variable to true with prob. $1/2$. \\
\end{alg}
Now, $\mathbf{Pr}[C \text{ sat. by assignment from algo.}] \geq 1-\frac{1}{2^k}$ and thus $\mathbf{E}[\text{\# clauses sat.}] \geq m(1-2^{-k})$, and the theorem follows by THM.

\paragraph{Method 1} The above is a $1/2$ approximation, i.e. $\geq m(1-2^{-k})$ clauses in expectation where each clause has $\geq k$ literals.

\paragraph{Method 2} Is a $\beta_k$-approximation. Let $C_i^-$, $C_i^+$ be the sets of negated and remaining variables in clause $i$, respectively. ILP formulation:
\begin{align*}
\max & \sum_{\text{clause } i} z_i\\
\text{s.t. }& \sum_{j \in C^+_i} y_j + \sum_{j \in C_i^-}(1-y_j) \geq z_i \; \forall \text{ clauses } i \\
\text{where }& y_j,z_i \in \left\{ 0,1 \right\} \; \forall i,j
\end{align*}

\begin{alg}
Solve linear relaxation of ILP where $\hat{y}_j, \hat{z}_i$ are fractional solutions and $\overline{y}_j$ is the rounded solution, s.t. $\mathbf{Pr}[\overline{y}_j = 1] = \hat{y}_j$. \\
\end{alg}

Consider a clause $i$ with $k$ literals. Assume that $C_i^- = \emptyset$. Then
\begin{align*}
\mathbf{\text{clause } i \text{ sat.}} &=1-\prod_{j \in C_i^+}(1-\hat{y}_j) \\
&\geq 1-\prod_{j \in C_i^+}\left(1-\frac{\hat{z}_j}{k}\right)=1-\left(1-\frac{\hat{z}_j}{k}\right)^k \\
&\geq\beta_k\hat{z}_i\text{, where } \beta_k=1-\left(1-\frac{1}{k}\right)^k
\end{align*}
Then $\mathbf{Pr}[\text{\# clauses sat. assuming every clause has at least } k \text{ literals}] \geq \sum_{\text{clause } i}\beta_k \cdot OPTLP \geq \beta_k \cdot OPT$. Thus $\sum \hat{z}_i \geq OPT$. Now, let $n_i = \mathbf{E}[\text{\# clauses sat. by method } i]$. Then \\
\begin{thm}
$\max \left\{ n_1, n_2 \right\} \geq \frac{3}{4} \cdot OPT$
\paragraph{Proof:}
\begin{align*}
\max \left\{ n_1, n_2 \right\} &\geq \frac{n1+n2}{2} \geq \sum_k \sum_{i \in S_k} \frac{(1-2^{-k}) + \beta_k}{2} \hat{z}_i \\
&\geq \sum_k \sum_{i \in S_k} \frac{3}{4} \hat{z}_i = \frac{3}{4} \sum_{i \in S_k}  \hat{z}_i \geq \frac{3}{4} \cdot OPT
\end{align*}
\end{thm}

\subsection*{OR-concentrator}
\begin{dfn}
An $(n,d,\alpha,c)$ OR-concentrator is a bipartite graph $G=(L,R,E)$ with $|L|=|R|=n$ s.t.
\begin{enumerate}
\item Every vertex in $L$ has deg. $\leq d$ and 
\item $\forall S \subseteq L : |S| \leq \alpha n \Rightarrow |N(S)| \geq c |S|$ \\
\end{enumerate}
\end{dfn}

\begin{thm}
Given $n \in \mathbb{N}, \; \exists (n,18,\frac{1}{3},2)$ OR-conc.
\paragraph{Proof:} For each $v \in L$, rand. algo picks $d$ neighbours in $R$, uniformly at random and with replacement. Connect $v$ with edges to its chosen neighbours.

Let $\varepsilon_S : \exists S \subseteq L, |S|=s, |N(S)| \leq cs$. Then
\begin{align*}
\mathbf{Pr}[\varepsilon_S]&=\binom{n}{s}\binom{n}{cs}\binom{cs}{n}^{ds} \\
&\leq \binom{ne}{s}^s\binom{ne}{cs}^cs\binom{cs}{n}^{ds} \\
&= \left( \binom{ne}{s}\binom{ne}{2s}^2\binom{2s}{n}^{18} \right)^s  \\
&= \left( n^{-15} s^{15} e^3 2^{16} \right)^5 \\
&\leq \left( n^{-15} \left( \frac{1}{3^n} \right)^{15} e^3 2^{16} \right)^5 \\
&= \left( \left( \frac{1}{3} \right)^{-15} e^3 2^{16} \right)^5 < \left( \frac{1}{2} \right)^s
\end{align*}
And $\mathbf{Pr}\left[ \cup_{S=1}^{\alpha n} \varepsilon_S \right] \leq \sum_{S=1}^{\alpha n} \left(\frac{1}{2}\right)^S < 1$. By TPM, an OR-concentrator exists with positive probability.
\end{thm}

\subsection*{Probability amplification}
Given an algorithm $A(x,t)$ where if $x \not\in L$ then $A(x,r) = 0$ $\forall r$ and if $x \in L$ then $A(x,r)=1$ for at least half the choices of $r$. \\

\begin{thm}
For suff. large $n \in \mathbb{N}$, there exists a bipartite graph $G=(L,R,E)$ s.t. $|L| = n$, $|R| = 2^{\log^2 n}$ and
\begin{enumerate}
\item every vertex in $R$ has deg. $\leq 12 \log^2 n$
\item $\forall S \subseteq L, |S|=\frac{n}{2} \Rightarrow |N(S)| \geq 2^{\log^2 n} - n$
\end{enumerate}
\paragraph{Proof: } For each $u \in L$, rand. algo picks $d=\frac{2^{\log^2 n}(4 \log^2 n)}{n}$ neighbours in $R$ uniformly and independently at random w. replacement. This gives a bipartite graph $G=(L,R,E)$. Let $v \in R$. We wish to apply Chernoff bound, so $\mathbf{E}[\deg(v)]=nd\frac{1}{2^{log^2 n}} = 2 \log^2 n$ and
\begin{align*}
\mathbf{Pr}\left[ \deg(v) > 3 \cdot 3 \log^2 n \right] < \left( \frac{e}{1+\delta}\right)^{1+\delta}\mu=\left( \frac{e}{3} \right)^{12\log^2 n}
\end{align*}
Taking the union bound over the events yields $$\mathbf{Pr}\left[ \bigcup_{v \in R} \left\{ \deg(v) > 12 \log^2 n \right\} \right] < 2^{\log^2 n} \cdot \left(\frac{e}{3} \right)^{12 \log^2 n} \leq \frac{1}{2}$$
Showing that the probability that condition 1 fails is at most 1/2. For condition 2:
\begin{align*}
\mathbf{Pr}[\text{cond. 2 fails}] &\leq \binom{n}{n/2}\left( \frac{2^{\log^2 n}}{n} \right)\left( 1- \frac{n}{2^{\log^2 n}} \right)^{\frac{dn}{2}} \\
&\leq \left( \frac{ne}{n/2}^{n/2} \right) \left( \frac{2^{\log^2 e}}{n} \right)^n e^{-\frac{dn}{2}\frac{n}{2^{\log^2 n}}} \\
&= (2e)^{n/2}\left( \frac{2^{\log^2 n}e}{n} \right)^n e^{-2n \log^2 n} \\
&< \frac{1}{2}
\end{align*}
Finally, by union bound we get $\mathbf{Pr}[\text{cond. 1 fails or cond. 2 fails}] < \frac{1}{2} + \frac{1}{2} = 1$. Theorem follows from TPM.

\end{thm}
To amplify prob, we can, with $\log^2 n$ random bits, pick a vertex $v \in R$ uniformly at random. For each $u \in N(v)$, run $A(x,r_u)$ where $r_u$ is the index of $u$. Output $\lor_{u \in N(v)} A(x,r_u)$. This yields a failure prob. $\leq \frac{n}{2^{\log^2 n}}$

\end{document}

### 2017-09-07 assignments/6/paper.tex
Added 5 + 6

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[]{algorithm2e}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{alg}[thm]{Algorithm}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Randomized Algorithms \\
       Assignment 6}

\author{Viktor Hansen \& Simon Rueskov Schleicher}

\maketitle

\begin{abstract}
  This is the sixth weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{Exercise 7.4}
If we use the fingerprint function used for polynomial identities as F, we can solve the problem in the same way, but this takes more time.\\
With the fingerprint function described in the chapter, when we have calculated $F(X(j))$ we can calculate $F(X(j+1))$ as $F(X(j+1))=(F(X(j))-x_j*2^m)*2+x_{m+1}\text{ mod }p$, which can be done in $O(1)$ time. The reason we can do this is that if we call the m-bit integer starting at $x_j$ for $Z_j$ and the m-bit integer starting at $x_{j+1}$ for $Z_{j+1}$, then to change $Z_j$ into $Z_{j+1}$ we need to remove the first bit (by subtracting $x_j*2^m$), shift everything 1 place to the left (by multiplying with $2$) and add $x_{m+1}$. And calculating this modulo p doesn't change the result.\\
With the fingerprint function used for polynomial identities on the other hand, all of the bits are multiplied by something different whether we're calculating $F(X(j))$ or $F(X(j+1))$, and there's no easy conversion between this. So we have to calculate for all m bits each time thus taking $O(m)$ time to get from $F(X(j))$ to $F(X(j+1))$.

\section*{Problem 7.1}
Define the multivariate polynomial $Q(x_1,...,x_n)=ABr-Cr$ where r is the vector consisting of $x_1,...,x_n$ in that order. We have $AB=C\Rightarrow ABr=Cr\Rightarrow ABr-Cr=0$ for any $r$, so we get $Q\equiv0$ if $AB=C$. If $AB\neq C$ we get $AB-C\neq0$, so there exists an $r$, where $ABr-Cr\neq0$. We therefore get $Q\equiv 0$ if and only if $AB=C$.\\
We choose the set $S=\{0,1\}$, and get $|S|=2$. According to Theorem 7.2 (Schwartz-Zippel Theorem), if we let $r_1,...,r_n$ be chosen independently and uniformly at random from $S$, then $Pr[Q(r_1,...,r_n)=0|Q(x_1,...,x_n)\not\equiv0]\leq\frac{d}{|S|}=\frac{d}{2}$. With $d=1$ we get $Pr[Q(r_1,...,r_n)=0|Q(x_1,...,x_n)\not\equiv0]\leq\frac{1}{2}$.\\
$AB\neq C$ is equivalent to $Q(x_1,...,x_n)\not\equiv0$. So if we assume $AB\neq C$, we get $Pr[Q(r_1,...,r_n)=0]\leq\frac{1}{2}$. From our definition of $Q$ we get $Q(r_1,...,r_n)=0\Leftrightarrow ABr-Cr=0\Leftrightarrow ABr=Cr$, so we get $Pr[Q(r_1,...,r_n)=0]\leq\frac{1}{2}\Leftrightarrow Pr[ABr=Cr]\leq\frac{1}{2}$, which is what we wanted to show.

\section*{Problem 7.4}
We call the integers in $S_1$ $s_1,...,s_n$ and the integers in $S_2$ $s_{n+1},...,s_{2n}$. We say that the lowest value of the integers in $S_1$ and $S_2$ is $m_1$ and the highest value of the integers in $S_1$ and $S_2$ is $m_2$. We define $m=m_2-m_1$. We can find $m_1$ and $m_2$ in $O(n)$ time.\\
A way to solve the problem involving a verification of a polynomial identity, is by defining $Q(x_{m_1},...,x_{m_2})=\Sigma_{i=1}^ns_ix_{s_i}-\Sigma_{i=n+1}^{2n}s_ix_{s_i}$. A calculation of $Q$ takes $O(n+m)$ time and requires $O(m)$ space.\\
We need to show that if $S_1$ and $S_2$ are identical in the sense that each integer occurs the same number of times in both sets, then $Q\equiv0$. We can sort each of the sets individually without changing the result of $Q$, and then we see $\Sigma_{i=1}^ns_ix_{s_i}-\Sigma_{i=n+1}^{2n}s_ix_{s_i}=\Sigma_{i=0}^ns_ix_{s_i}-s_{i+n}x_{s_{i+n}}=\Sigma_{i=0}^ns_ix_{s_i}-s_ix_{s_i}=0$. We also need to show that if $S_1$ and $S_2$ are not identical in the sense that each integer occurs the same number of times in both sets, then $Q\not\equiv0$. In this case there has to be an integer that occurs more times in $S_1$ than in $S_2$. Call this integer j. We get $Q(0,...,x_j=1,...,0)>0$, so $Q\not\equiv0$.\\
Now we have $Q\equiv0$ if and only if $S_1$ and $S_2$ are identical in the sense that each integer occurs the same number of times in both sets, and we can use Theorem 7.2 to solve the problem.\\
Comparing this solution with the deterministic solution given in the book, we see that if the integers in the sets has a small range ($m$ is small), then this solution is faster. But for large ranges of integers ($m$ is large) this solution is both slower and uses a lot of space.

\section*{Summary}
If we have 3 matrices $A$, $B$ and $C$, and we want to verify that $AB=C$, then we can use Theorem 7.1:\\
\textbf{Theorem 7.1: } Let $A$, $B$ and $C$ be $n*n$ matrices over $F$ such that $AB\neq C$. Then for r chosen uniformly at random from $\{0,1\}^n$, $Pr[ABr=Cr]\leq\frac{1}{2}$.\\
We of course also have that if $AB=C$, then $ABr=Cr$.\\
\textbf{Theorem 7.2 (Schwartz-Zippel Theorem): } Let $Q(x_1,...,x_n)\in F[x_1,...,x_n]$ be a multivariate polynomial of total degree d. Fix any finite set $S\subseteq F$, and let $r_1,...,r_n$ be chosen independently and uniformly at random from $S$. Then $$Pr[Q(r_1,...,r_n)=0|Q(x_1,...,x_n)\not\equiv0]\leq\frac{d}{|S|}.$$
Perfect Matchings in Graphs:\\
\textbf{Theorem 7.3 (Edmonds' Theorem): } Let $A$ be the $n*n$ matrix obtained from $G(U,V,E)$ as follows: $A_{ij}=x_{ij}$ if $(u_i,v_j)\in E$ and $A_{ij}=0$ otherwise. Define the multivariate polynomial $Q(x_{11},x_{12},...,x_{nn})$ as being equal to det($A$). Then, G has a perfect matching if and only if $Q\not\equiv0$.\\
Verifying Equality of Strings:\\
Interpret the data as $n$-bit integers $a$ and $b$, by defining $a=\Sigma_{i=1}^na_i2^{i-1}$ and $b=\Sigma_{i=1}^nb_i2^{i-1}$. Define the fingerprint function $F_p(x)=x\text{ mod }p$ for a prime p. For any number k, let $\pi(k)$ be the number of distinct primes less than k. Choose a $\tau=tn\log tn$ for large t. Then we have the following Theorem:\\
\textbf{Theorem 7.5: } $Pr[F_p(a)=F_p(b)|a\neq b]\leq\frac{n}{\pi(\tau)}=O(\frac{1}{t})$.\\
Pattern matching:\\
A text is a string $X=x_1x_2...x_n$ and a pattern is a string $Y=y_1y_2...y_m$, both over a fixed finite alphabet $\Sigma$, such that $m\leq n$. The pattern matching problem is to find out if m occurs anywhere in n. The simple brute force deterministic algorithm takes $O(nm)$ time. We can make a Monte Carlo algorithm to solve this problem:\\
For any string $Z\in\{0,1\}^m$, interpret $Z$ as an $m$-bit integer and define $F_p(Z)=Z\text{ mod }p$. Assume that $p$ is chosen uniformly at random from the set of primes smaller than a threshold $\tau$. We get $Pr[F_p(Y)=F_p(X(j))|Y\neq X(j)]\leq\frac{m}{\pi(\tau)}=O\left(\frac{m\log\tau}{\tau}\right).$. $Pr[\text{a false match occurs}]=O(\frac{1}{n})$.\\
\textbf{Theorem 7.6: } The Monte Carlo algorithm for pattern matching requires $O(n+m)$ time and has a probability of error $O(1/n)$.

\end{document}
